---
title: "Coevolution with a seed bank: population density"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
library(renv)
# # Initialize lock file when first setting repo
# renv::init()

library(here)
# # set root at project file ("coevolution-ts"). 
# # Done once at initial setup.
# here::set_here()

library(tidyverse)
library(cowplot)

# save the state of the project library to the lockfile (called renv.lock),
renv::snapshot()

# save default par for base graphics
par.default <- par()
```

## Population density data  
I have previously consolidated all the data on population density collected for this experiment. This includes:

| Population | What it is | Method of measurment |
| - | - | - |
|spore / mL | host spore density | flow-cytometry: low SYBR population |
|veg / mL | host vegetative cell density | flow-cytometry: high SYBR population |
| cell /mL | total host cell density (spore+veg) | flow-cytometry
| phage/ mL | total SPO1 phage density (free phage) | qPCR (SYBR)  with phage specific primers and lysate serial dilution as standard | 

```{r load data, include=FALSE}
d <- read_csv(here("data","coevolution_pop_density.csv"))
```

### Decomposition of the time series  
From QB handout on temporal diversity: 

> A time series can be broken down into different "categories", which can provide insight into factors contributing to temporal patterns.   
The common categories of a time series decomposition are the trend (i.e., increasing or decreasing value in the series), seasonality (repeating short-term cycles), and residual error (i.e., statistical noise due to stochastic changes or sampling error).  

```{r, include=FALSE}
# initailise tibble for collecting decompsed data
d.decomp <- 
  tibble(Time=NA,
         line=NA,
         culture=NA,
         pop=NA,
         observed= NA,
         trend= NA,
         seasonal= NA,
         random= NA)


for (ln in levels(as.factor(d$line))){
  for (cult in levels(as.factor(d$culture))){
    for (pop in c("phage.ml","spore.ml","veg.ml","cell.ml")){
      
  #not all the lines have measurments of all the populations:
        #Ct lines have no phage
        #SN lines have no spores
  # the loop skips those combinations
      if (grepl("SN",cult) & pop=="spore.ml")
        next
      
      if (grepl("Ct",cult) & pop=="phage.ml")
        next
      
  tmp <- 
    d%>%
  filter(Time>=0)%>%
  filter(culture==cult)%>%
  filter(line==ln)%>%
  select(all_of(pop))%>%
ts(.[,pop], start=0, end=14,frequency = 2 )%>%
  decompose()
      
  # get trend data
  d.decomp <- 
  tibble(Time=seq(0,14,0.5),
         line=ln,
         culture=cult,
         pop=pop,
         observed= as.vector(tmp$x),
         trend= as.vector(tmp$trend),
         seasonal= as.vector(tmp$seasonal),
         random= as.vector(tmp$random))%>%
    bind_rows(d.decomp,.)
  tmp$trend%>%as.vector()
  
    }

  }

}

#remove first row used to initialise
d.decomp <- d.decomp[-1,]

#delete temporary variables
rm(tmp,cult,ln,pop)

```


### Plot decomposed trend data

```{r}
d.decomp%>%
  filter(!is.na(trend))%>%
  filter(pop!="cell.ml")%>%
  ggplot(aes(x=Time, y=trend))+
  geom_line(aes(color=pop), size=1)+
  facet_grid(line~culture)+
  theme_cowplot()+
  panel_border()+
  scale_y_log10()+
  theme(legend.position = "bottom")
  
```

```{r}
d.decomp%>%
  filter(!is.na(trend))%>%
  filter(pop=="cell.ml" | pop=="phage.ml")%>%
  ggplot(aes(x=Time, y=trend))+
  geom_line(aes(color=pop), size=1)+
  facet_grid(line~culture)+
  theme_cowplot()+
  panel_border()+
  scale_y_log10()+
  theme(legend.position = "bottom")
  
```

Plot by treatment
```{r}
d.decomp%>%
  #add columns on phage treatment and seed bank
  mutate(phage=case_when(grepl("O", culture) ~ "SPO1",
                         grepl("Ct", culture) ~ "no_Phage",
                         grepl("ANC", line) ~ "no_Phage",
                         grepl("blank", culture) ~ "blank"))%>%
  mutate(seed.bank=case_when(grepl("SN", culture) ~ "none",
                              grepl("WS", culture) ~ "short",
                              grepl("WL", culture) ~ "long"))%>%
   mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))%>%
  #add flask to be used as grouping variable
  mutate(flask=paste(culture,line, sep = "-"))%>%
  #remove edges fr which trend is NA
  filter(!is.na(trend))%>%
  # looking only at total cells
  filter(pop=="cell.ml")%>%
  ggplot(aes(x=Time, y=trend))+
  geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
  facet_wrap(~phage)+
  theme_cowplot()+
  panel_border()+
  scale_y_log10()+
  theme(legend.position = "bottom")+
  ylab("cell/ml (trend)")
  
```

# Phase Plane w. trend data

```{r, echo=FALSE}
# Phase plane plots
d.decomp %>%
   filter(Time>=0)%>%
select(Time,culture,line,pop,trend)%>%
  filter(pop=="cell.ml" | pop=="phage.ml")%>%
  pivot_wider(values_from = trend, names_from=pop)%>%
  filter(!is.na(phage.ml))%>%
ggplot( aes(x=log10(cell.ml), y=log10(phage.ml) ,colour=Time)) +
  geom_path(arrow = arrow(type = "closed", angle = 15, length = unit(0.1, "inches"))) +
  facet_grid(line~culture)+#, scales = "free") +
  xlab("Host (log10)") + ylab("Phage (log10)") +
  theme_cowplot()+
   panel_border()+
   scale_color_viridis_c()+
   ggtitle("Decomposed trend")

```

# animate
```{r, include=FALSE}
library(gganimate)
library(gifski)


# Phase plane plots

p.anim <- 
d.decomp %>%
   filter(Time>=0)%>%
select(Time,culture,line,pop,trend)%>%
  filter(pop=="cell.ml" | pop=="phage.ml")%>%
  pivot_wider(values_from = trend, names_from=pop)%>%
  filter(!is.na(phage.ml))%>%
ggplot( aes(x=log10(cell.ml), y=log10(phage.ml) ,colour=Time)) +
  geom_path(arrow = arrow(type = "closed", angle = 15, length = unit(0.1, "inches"))) +
  facet_grid(line~culture, scales = "free") +
  xlab("Host (log10)") + ylab("Phage (log10)") +
  theme_cowplot()+
   panel_border()+
   scale_color_viridis_c()+
   transition_reveal(Time)+
   ggtitle(label = "All time points", subtitle = "Time {frame_along}")

animate(p.anim, duration = 15, fps = 10,  renderer =gifski_renderer())
anim_save(here("plots","phase_plane_all.gif"))
```

# Repeated-Measures Analysis-of-Variance (RM-ANOVA)

*Following QB module*

