---
title: "Coevolution with a seed bank: population density"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
library(renv)
# install to projectlibray the resuired packages
renv::restore()
# # Initialize lock file when first setting repo
# renv::init()

library(here)
# # set root at project file ("coevolution-ts"). 
# # Done once at initial setup.
# here::set_here()

library(tidyverse)
library(cowplot)

library(nlme)
library(pander)
library(emmeans)
library(codyn)

# save the state of the project library to the lockfile (called renv.lock),
renv::snapshot()

# save default par for base graphics
par.default <- par()
```

## Population density data  
I have previously consolidated all the data on population density collected for this experiment. This includes:

| Population | What it is | Method of measurement |
| - | - | - |
|spore / mL | host spore density | flow-cytometry: low SYBR population |
|veg / mL | host vegetative cell density | flow-cytometry: high SYBR population |
| cell /mL | total host cell density (spore+veg) | flow-cytometry
| phage/ mL | total SPO1 phage density (free phage) | qPCR (SYBR)  with phage specific primers and lysate serial dilution as standard | 

```{r load data, echo=FALSE}
d <- read_csv(here("data","coevolution_pop_density.csv"))

#add columns on host strain and phage treatment  as encoded in culture name
d <- 
  d%>%
   mutate(host=case_when(grepl("W", culture) ~ "WT",
                         grepl("SN", culture) ~ "dSpoIIE",
                         grepl("dS", culture) ~ "dSpoIIE"))%>%
  
  mutate(phage=case_when(grepl("O", culture) ~ "SPO1",
                         grepl("Ct", culture) ~ "no_Phage",
                         grepl("ANC", line) ~ "no_Phage"))%>%

   mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))%>%
  #add flask to be used as grouping variable
  mutate(flask=paste(culture,line, sep = "-"))

# plot total cells of raw data
d%>%
  filter(Time>=0)%>%
    ggplot(aes(x=Time, y=cell.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      facet_wrap(~phage)+
      theme_cowplot()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom")+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 1e9))
```

### Decomposition of the time series  
From QB handout on temporal diversity: 

> A time series can be broken down into different "categories", which can provide insight into factors contributing to temporal patterns.   
The common categories of a time series decomposition are the trend (i.e., increasing or decreasing value in the series), seasonality (repeating short-term cycles), and residual error (i.e., statistical noise due to stochastic changes or sampling error).  

First let's look at decomposition of a single sample (WSO-L1), looking at total cells (log-transformed)

```{r, echo=FALSE}

    d%>%
  filter(Time>=0)%>%
  filter(culture=="WSO")%>%
  filter(line=="L1")%>%
  mutate(cell.ml=log10(cell.ml))%>%
  select(cell.ml)%>%
ts(.[,"cell.ml"], start=0, end=14,frequency = 2 )%>%
  decompose()%>%
  plot()

```

The decomposition removes the bi-daily fluctuations (season component) and reveals 2 population cycles in the first week, followed by a steady rise in the second week.

Next let's decompose in the same manner all of the data.

```{r, include=FALSE}
# initailise tibble for collecting decompsed data
d.decomp <- 
  tibble(Time=NA,
         line=NA,
         culture=NA,
         pop=NA,
         observed= NA,
         trend= NA,
         seasonal= NA,
         random= NA)


for (ln in levels(as.factor(d$line))){
  for (cult in levels(as.factor(d$culture))){
    for (pop in c("phage.ml","spore.ml","veg.ml","cell.ml")){
      
  #not all the lines have measurments of all the populations:
        #Ct lines have no phage
        #SN lines have no spores
  # the loop skips those combinations
      if (grepl("SN",cult) & pop=="spore.ml")
        next
      
      if (grepl("Ct",cult) & pop=="phage.ml")
        next
      
  tmp <- 
    d%>%
  filter(Time>=0)%>%
  filter(culture==cult)%>%
  filter(line==ln)%>%
  select(all_of(pop))%>%
ts(.[,pop], start=0, end=14,frequency = 2 )%>%
  decompose()
      
  # get trend data
  d.decomp <- 
  tibble(Time=seq(0,14,0.5),
         line=ln,
         culture=cult,
         pop=pop,
         observed= as.vector(tmp$x),
         trend= as.vector(tmp$trend),
         seasonal= as.vector(tmp$seasonal),
         random= as.vector(tmp$random))%>%
    bind_rows(d.decomp,.)
  tmp$trend%>%as.vector()
  
    }

  }

}

#remove first row used to initialise
d.decomp <- d.decomp[-1,]

#delete temporary variables
rm(tmp,cult,ln,pop)

#add columns on host strain,phage treatment and seed bank as encoded in culture name
d.decomp <- 
   d.decomp%>%
   mutate(host=case_when(grepl("W", culture) ~ "WT",
                         grepl("SN", culture) ~ "dSpoIIE",
                         grepl("dS", culture) ~ "dSpoIIE"))%>%
  
  mutate(phage=case_when(grepl("O", culture) ~ "SPO1",
                         grepl("Ct", culture) ~ "no_Phage",
                         grepl("ANC", line) ~ "no_Phage"))%>%
  
  mutate(seed.bank=case_when(grepl("SN", culture) ~ "none",
                              grepl("WS", culture) ~ "short",
                              grepl("WL", culture) ~ "long"))%>%
   mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))%>%
  #add flask to be used as grouping variable
  mutate(flask=paste(culture,line, sep = "-"))

```


### Plot decomposed trend data

Plot by treatment, as above.  
```{r,echo=FALSE}
d.decomp%>%
  #remove edges in which trend is NA
  filter(!is.na(trend))%>%
  # looking only at total cells
  filter(pop=="cell.ml")%>%
  ggplot(aes(x=Time, y=trend))+
  geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
  facet_wrap(~phage)+
  theme_cowplot()+
  panel_border()+
  scale_y_log10()+
  theme(legend.position = "bottom")+
  ylab("cell/ml (trend)")
  
```

The patterns are somewhat clearer. Especially obvious is the differences between WT (long and short seed bank) and SPOIIE (no seed bank) in the first week when infected with phage. The 3 lines with the non-sporulating mutant all have similar looking cycles. 

And now adding to total cells also the phages.  

```{r}
d.decomp%>%
  filter(!is.na(trend))%>%
  filter(pop=="cell.ml" | pop=="phage.ml")%>%
  ggplot(aes(x=Time, y=trend))+
  geom_line(aes(color=pop), size=1)+
  facet_grid(line~culture)+
  theme_cowplot()+
  panel_border()+
  scale_y_log10()+
  theme(legend.position = "bottom")
```

Next plotting all the decomposed trend data: spores, veg and phages. 
```{r}
d.decomp%>%
  filter(!is.na(trend))%>%
  filter(pop!="cell.ml")%>%
  ggplot(aes(x=Time, y=trend))+
  geom_line(aes(color=pop), size=1)+
  facet_grid(line~culture)+
  theme_cowplot()+
  panel_border()+
  scale_y_log10()+
  theme(legend.position = "bottom")
  
```


Finally pulling together spore and veg data from the 3 seed bank treatments:

```{r}
d.decomp%>%
  #remove edges in which trend is NA
  filter(!is.na(trend))%>%
  # looking only at cells by type
  filter(pop=="spore.ml" | pop=="veg.ml")%>%
  ggplot(aes(x=Time, y=trend))+
  geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
  facet_grid(pop~phage)+
  theme_cowplot()+
  panel_border()+
  scale_y_log10()+
  theme(legend.position = "bottom")+
  ylab("cell/ml (trend)")
  
```


# Stability

From QB:

> The stability of an aggregate measure of an ecological system can be assessed by measuring its variability.
One way to characterize variability is the Coefficient of Variation (CV). 
The CV relativizes the standard deviation of a variable to its mean value because variance scales with the mean.
By using the CV, we can more easily compare the variability of systems with different mean values.
We can calculate the CV as follows: 
$$CV = \frac{\sigma}{\mu}$$
where $\sigma$ is the standard deviation and $\mu$ is the mean value. 
Higher CV indicates more variability, and lower CV indicates less variability. 
Therefore, we can measure stability as the inverse of CV:
$$\textrm{Stability} = \frac{1}{CV}$$

```{r, echo=F}

d.cell.stab <- 
  d.decomp%>%
  #remove edges in which trend is NA
  filter(!is.na(trend))%>%
  # looking only at total cells
  filter(pop=="cell.ml")%>%
  as.data.frame()

cell.stab <- 
  community_stability(df = d.cell.stab,
                    time.var = "Time",
                    abundance.var = "trend",
                    replicate.var = "flask")

cell.stab <- 
  cell.stab%>%
    mutate(host=case_when(grepl("W", flask) ~ "WT",
                         grepl("SN", flask) ~ "dSpoIIE",
                         grepl("dS", flask) ~ "dSpoIIE"))%>%
  
    mutate(phage=case_when(grepl("O", flask) ~ "SPO1",
                           grepl("Ct", flask) ~ "no_Phage"))%>%
    
    mutate(seed.bank=case_when(grepl("SN", flask) ~ "none",
                                grepl("WS", flask) ~ "short",
                                grepl("WL", flask) ~ "long"))%>%
     mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))%>%
    
    mutate(line=str_sub(flask, -2,-1))

#summarise
sum.cell.stab <- 
  cell.stab%>%
    group_by(seed.bank,phage) %>% 
    summarise(
      sd = sd(stability),
      stability = mean(stability))
    
    
  # plot
cell.stab%>%
  ggplot(aes(x=seed.bank, y=stability))+
  geom_crossbar(data=sum.cell.stab,
                  aes(ymin=stability-sd, ymax=stability+sd), width=0.3)+
  geom_dotplot(binaxis='y', stackdir='center', fill="white", alpha=0.5)+
  # geom_jitter(shape=21, fill="grey", size=2, width = 0.05, height = 0)+
  facet_wrap(~phage, scales = "free")+
  theme_cowplot()+
  ggtitle("decomposed trend data on total cell density")
```

In the presence of phage, the presence of a seed bank stabilizes the host population!!

Now the same analysis with the observed data

```{r, echo=F}

d.cell.stab <- 
  d.decomp%>%
  #remove edges in which trend is NA
  filter(!is.na(observed))%>%
  # looking only at total cells
  filter(pop=="cell.ml")%>%
  as.data.frame()

cell.stab <- 
  community_stability(df = d.cell.stab,
                    time.var = "Time",
                    abundance.var = "observed",
                    replicate.var = "flask")

cell.stab <- 
  cell.stab%>%
    mutate(host=case_when(grepl("W", flask) ~ "WT",
                         grepl("SN", flask) ~ "dSpoIIE",
                         grepl("dS", flask) ~ "dSpoIIE"))%>%
  
    mutate(phage=case_when(grepl("O", flask) ~ "SPO1",
                           grepl("Ct", flask) ~ "no_Phage"))%>%
    
    mutate(seed.bank=case_when(grepl("SN", flask) ~ "none",
                                grepl("WS", flask) ~ "short",
                                grepl("WL", flask) ~ "long"))%>%
     mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))%>%
    
    mutate(line=str_sub(flask, -2,-1))

#summarise
sum.cell.stab <- 
  cell.stab%>%
    group_by(seed.bank,phage) %>% 
    summarise(
      sd = sd(stability),
      stability = mean(stability))
    
    
  # plot
cell.stab%>%
  ggplot(aes(x=seed.bank, y=stability))+
  geom_crossbar(data=sum.cell.stab,
                  aes(ymin=stability-sd, ymax=stability+sd), width=0.3)+
  geom_dotplot(binaxis='y', stackdir='center', fill="white", alpha=0.5)+
  # geom_jitter(shape=21, fill="grey", size=2, width = 0.05, height = 0)+
  facet_wrap(~phage, scales = "free")+
  theme_cowplot()+
  ggtitle("Observed data on total cell density")
```

Exactly the same. Is there statistical support for seed bank stability?

```{r stability stats}
aov_stab <- 
  aov(stability~seed.bank, 
    data = filter(cell.stab, phage=="SPO1"))
  
summary(aov_stab)
```

<<<<<<< HEAD
Yes, there is! Post-hoc:

```{r}
TukeyHSD(aov_stab)
```

```{r}
TukeyHSD(aov_stab)%>%plot(col="blue")
```

The difference in stability is driven by a significant difference in stability between the long-residence seed-bank and the no seed-bank treatments. the short-residence seed-bank is not significantly distinguishable from either of the other treatments.


# Repeated-Measures Analysis-of-Variance (RM-ANOVA)

*Following QB module*

## RM-ANOVA on decomposed trend total cell data  
```{r,  message = FALSE, warning = FALSE}

cell.rm <- 
  d.decomp%>%
  #remove edges in which trend is NA
  filter(!is.na(trend))%>%
  # looking only at total cells
  filter(pop=="cell.ml")%>%
  lme(trend ~ phage * seed.bank, random = ~ 1 | line, 
            correlation = corAR1(form = ~ 1 | line),
            data = .)

# Look at detailed output
summary(cell.rm) 

# Obtain F-test
anova(cell.rm)

# # Make cleaner ANOVA table
# set.caption("RMANOVA for Portal")
# pander(anova(cell.rm))

# # Use `lsmeans` package for time-corrected marginal means
# lsmeans(cell.rm, ~seed.bank+phage)  %>% plot()+theme_cowplot()
```

## RM-ANOVA on measured total cell data 
```{r,  message = FALSE, warning = FALSE}

cell.rm.obs <- 
  d%>%
  #remove edges in which trend is NA
  filter(!is.na(trend))%>%
  # looking only at total cells
  filter(pop=="cell.ml")%>%
  lme(trend ~ phage * seed.bank, random = ~ 1 | line, 
            correlation = corAR1(form = ~ 1 | line),
            data = .)

# Look at detailed output
summary(cell.rm) 

# Obtain F-test
anova(cell.rm)

# # Make cleaner ANOVA table
# set.caption("RMANOVA for Portal")
# pander(anova(cell.rm))

# # Use `lsmeans` package for time-corrected marginal means
# lsmeans(cell.rm, ~seed.bank+phage)  %>% plot()+theme_cowplot()
```

## RM-ANOVA on decomposed trend phage data  
```{r,  message = FALSE, warning = FALSE}

phage.rm <- 
  d.decomp%>%
  # looking only at phages
  filter(pop=="phage.ml")%>%
  #remove edges in which trend is NA 
  # AND remove no-phage controls
  filter(!is.na(trend))%>%
  lme(trend ~ seed.bank, random = ~ 1 | line, 
            correlation = corAR1(form = ~ 1 | line),
            data = .)

# Look at detailed output
summary(phage.rm) 

# Obtain F-test
anova(phage.rm)

# # Make cleaner ANOVA table
# set.caption("RMANOVA for Portal")
# pander(anova(cell.rm))

# # Use `lsmeans` package for time-corrected marginal means
# lsmeans(phage.rm, ~seed.bank) %>% plot()+theme_cowplot()
```


# Phase Plane w. decomposed trend data

```{r, echo=FALSE}
# Phase plane plots
d.decomp %>%
   filter(Time>=0)%>%
select(Time,culture,line,pop,trend)%>%
  filter(pop=="cell.ml" | pop=="phage.ml")%>%
  pivot_wider(values_from = trend, names_from=pop)%>%
  filter(!is.na(phage.ml))%>%
ggplot( aes(x=log10(cell.ml), y=log10(phage.ml) ,colour=Time)) +
  geom_path(arrow = arrow(type = "closed", angle = 15, length = unit(0.1, "inches"))) +
  facet_grid(line~culture)+#, scales = "free") +
  xlab("Host (log10)") + ylab("Phage (log10)") +
  theme_cowplot()+
   panel_border()+
   scale_color_viridis_c()+
   ggtitle("Decomposed trend")

```


```{r animate, eval=FALSE}
library(gganimate)
library(gifski)


# Phase plane plots

p.anim <- 
d.decomp %>%
   filter(Time>=0)%>%
select(Time,culture,line,pop,trend)%>%
  filter(pop=="cell.ml" | pop=="phage.ml")%>%
  pivot_wider(values_from = trend, names_from=pop)%>%
  filter(!is.na(phage.ml))%>%
ggplot( aes(x=log10(cell.ml), y=log10(phage.ml) ,colour=Time)) +
  geom_path(arrow = arrow(type = "closed", angle = 15, length = unit(0.1, "inches"))) +
  facet_grid(line~culture, scales = "free") +
  xlab("Host (log10)") + ylab("Phage (log10)") +
  theme_cowplot()+
   panel_border()+
   scale_color_viridis_c()+
   transition_reveal(Time)+
   ggtitle(label = "All time points", subtitle = "Time {frame_along}")

animate(p.anim, duration = 15, fps = 10,  renderer =gifski_renderer())
anim_save(here("plots","phase_plane_all.gif"))
```
