---
title: "Coevolution with a seed bank: population density"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
library(renv)
# # Initialize lock file when first setting repo
# renv::init()

library(here)
# # set root at project file ("coevolution-ts"). 
# # Done once at initial setup.
# here::set_here()

library(tidyverse)
library(cowplot)

library(tsibble)
library(feasts)
# save the state of the project library to the lockfile (called renv.lock),
renv::snapshot()

# save default par for base graphics
par.default <- par()
```

## Population density data  
I have previously consolidated all the data on population density collected for this experiment. This includes:

| Population | What it is | Method of measurment |
| - | - | - |
|spore / mL | host spore density | flow-cytometry: low SYBR population |
|veg / mL | host vegetative cell density | flow-cytometry: high SYBR population |
| cell /mL | total host cell density (spore+veg) | flow-cytometry
| phage/ mL | total SPO1 phage density (free phage) | qPCR (SYBR)  with phage specific primers and lysate serial dilution as standard | 

```{r load data, include=FALSE}
d <- read_csv(here("data","coevolution_pop_density.csv"))
```

### Decomposition of the time series  
From QB handout on temporal diversity: 

> A time series can be broken down into different "categories", which can provide insight into factors contributing to temporal patterns.   
The common categories of a time series decomposition are the trend (i.e., increasing or decreasing value in the series), seasonality (repeating short-term cycles), and residual error (i.e., statistical noise due to stochastic changes or sampling error).  

```{r, include=FALSE}

for (ln in levels(as.factor(d$line))){
  for (cult in levels(as.factor(d$culture))){
    for (pop in c("phage.ml","spore.ml","veg.ml","cell.ml")){
      
  #not all the lines have measurments of all the populations:
        #Ct lines have no phage
        #SN lines have no spores
  # the loop skips those combinations
      if (grepl("SN",cult) & pop=="spore.ml")
        next
      
      if (grepl("Ct",cult) & pop=="phage.ml")
        next
      
      d%>%
  filter(Time>=0)%>%
  filter(culture==cult)%>%
  filter(line==ln)%>%
  select(all_of(pop))%>%
ts(.[,pop], start=0, end=14,frequency = 2 )%>%
  decompose()%>%
  plot()
      
    }

  }

}

```

