---
title: "RM-ANOVA for coevolution with a seed bank - Decomposed data"
output: html_notebook
editor_options: 
  chunk_output_type: inline
  code_folding: hide
---

```{r setup, include=FALSE}
library(renv)
# install to project library the required packages
renv::restore()
# # Initialize lock file when first setting repo
# renv::init()

library(here)
# # set root at project file ("coevolution-ts"). 
# # Done once at initial setup.
# here::set_here()

library(tidyverse)
library(cowplot)


library(nlme)

# save the state of the project library to the lockfile (called renv.lock),
renv::snapshot()

```

## Population density data  
I have previously consolidated all the data on population density collected for this experiment. This includes:

| Population | What it is | Method of measurement |
| - | - | - |
|spore / mL | host spore density | flow-cytometry: low SYBR population |
|veg / mL | host vegetative cell density | flow-cytometry: high SYBR population |
| cell /mL | total host cell density (spore+veg) | flow-cytometry
| phage/ mL | total SPO1 phage density (free phage) | qPCR (SYBR)  with phage specific primers and lysate serial dilution as standard | 

```{r load data, echo=FALSE, message=FALSE}
d <- read_csv(here("data","coevolution_pop_density.csv"), )
```
In the first few days of the experiment cell numbers dropped below detection limit in 2 of 3 SNO lines. Missing data cannot be analyzed in RM-ANOVA. I will assign these points the detection limit values.

Points with missing data: 
```{r}
d%>%
  filter(cell.ml<10)
```


---  

**Flow-cytometry detection limit**

The flow cytometer stop conditions were either $50\mu L$ or 50,000 events, whatever came first.  
In addition, to account for false asignments by noise I required that a gated population have at least 100 events to be taken into account. So the minimal detected population would have 100 cells in $50\mu L$ of analyzed liquid. At this poimt of the experiment we were analyzing x100 diluted samples.

```{r}
detection.limit <- 100 * (1000/50) * 100
```

100 events X $1000/50\mu L$ X100 = `r detection.limit` 

I will use $1X10^5$, half the detection limit dor missing value points
---




```{r}
d <- d%>%
  mutate(cell.ml=if_else(cell.ml>10, cell.ml,detection.limit/2))
```

I will re-run the RM-ANOVA analysis on the data decomposed to remove seasona (transfer) fluctuations.
```{r, include=FALSE}
# initailise tibble for collecting decompsed data
d.decomp <- 
  tibble(Time=NA,
         line=NA,
         culture=NA,
         pop=NA,
         observed= NA,
         trend= NA,
         seasonal= NA,
         random= NA)


for (ln in levels(as.factor(d$line))){
  for (cult in levels(as.factor(d$culture))){
    for (pop in c("phage.ml","spore.ml","veg.ml","cell.ml")){
      
  #not all the lines have measurments of all the populations:
        #Ct lines have no phage
        #SN lines have no spores
  # the loop skips those combinations
      if (grepl("SN",cult) & pop=="spore.ml")
        next
      
      if (grepl("Ct",cult) & pop=="phage.ml")
        next
      
  tmp <- 
    d%>%
  filter(Time>=0)%>%
  filter(culture==cult)%>%
  filter(line==ln)%>%
  select(all_of(pop))%>%
ts(.[,pop], start=0, end=14,frequency = 2 )%>%
  decompose()

  # get trend data
  d.decomp <- 
  tibble(Time=seq(0,14,0.5),
         line=ln,
         culture=cult,
         pop=pop,
         observed= as.vector(tmp$x),
         trend= as.vector(tmp$trend),
         seasonal= as.vector(tmp$seasonal),
         random= as.vector(tmp$random))%>%
    bind_rows(d.decomp,.)
  tmp$trend%>%as.vector()
  
  #plot
    #   ggplot2::autoplot(tmp)+
    #     theme_bw()+
    # ggtitle(paste(cult,ln,pop))+
    #     ggsave(filename = here("plots","deomposition",paste(pop,cult,ln,".pdf", sep = "_")),width = 8, height = 6)
  
    }

  }

}

#remove first row used to initialise
d.decomp <- d.decomp[-1,]

#delete temporary variables
rm(tmp,cult,ln,pop)

# log transform data
d.decomp <- 
  d.decomp %>%
  filter(!is.na(trend))%>%
  mutate(trend=log10(trend))

#return to wide format
d.decomp.w <- 
  d.decomp%>%
  filter(!is.na(trend))%>%
  select(-observed,-seasonal,-random)%>%
  pivot_wider( names_from="pop", values_from =trend)

# #add columns on host strain,phage treatment and seed bank as encoded in culture name
d.decomp.w <-
   d.decomp.w%>%
   mutate(host=case_when(grepl("W", culture) ~ "WT",
                         grepl("SN", culture) ~ "dSpoIIE",
                         grepl("dS", culture) ~ "dSpoIIE"))%>%

  mutate(phage=case_when(grepl("O", culture) ~ "SPO1",
                         grepl("Ct", culture) ~ "no_Phage",
                         grepl("ANC", line) ~ "no_Phage"))%>%

  mutate(seed.bank=case_when(grepl("SN", culture) ~ "none",
                              grepl("WS", culture) ~ "short",
                              grepl("WL", culture) ~ "long"))%>%
   mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))%>%
  #add flask to be used as grouping variable
  mutate(flask=paste(culture,line, sep = "-"))%>%
  #remove setup data
  filter(Time>=0)

```




```{r}


# plot raw data
d.decomp.w%>%
    pivot_longer(cols=contains(".ml"), names_to="pop",values_to="num.ml")%>%
    ggplot(aes(x=Time, y=num.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      facet_grid(pop~phage)+
      theme_bw()+
      panel_border()+
      # scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(5, 10))
```



```{r}
d.decomp.w%>%
    ggplot(aes(x=Time, y=cell.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      # scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(5, log10(2e9)))

# plot total cells of  data
d.decomp.w%>%
  group_by(phage,seed.bank,Time)%>%
  summarise(n=n(),sd=sd(cell.ml), cell.ml=mean(cell.ml), sem=sd/sqrt(n))%>%
    ggplot(aes(x=Time, y=cell.ml))+
  geom_errorbar(aes(ymin=cell.ml-sem,ymax=cell.ml+sem,color=seed.bank), alpha=0.7)+
      geom_line(aes(color=seed.bank), size=1, alpha=0.7)+
      facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      # scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(5, log10(2e9)))

```


# Adjusting variable types for lme

In the lme models I will be using the fixd effects need to be secified as factore, in this experiment these ar: *Phage, seed.bank and time*. However for the autocorrelation specification time needs to be specified as an integer. To fulfill both requirements I will use for the time unit the experimental day rather than transfer, since we measured onc a day but twice per transfer. This will be simplty *time x 2*. I will then make fom that a separate varible which will be the factor of the time.

```{r}
d.4stat <- 
  d.decomp.w%>%
  # remove setup data
  filter(Time>=0)%>%
  #make time an integer for correlation structure
  # by converting to days as time unit
  mutate(Time=as.integer(2*Time))%>%
  # for the lme model all fixed effects need to be factors
  mutate(time.fct=as.factor(Time))%>%
  mutate(seed.bank=as.factor(seed.bank))%>%
  mutate(phage=as.factor(phage))
```


## Model selection  


### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r arma, message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the space in which the model conerges is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
    #this fllowing combination does not work: "iteration limit reached without convergence"
    if(P>=9 & Q==0) next
     if(P>=7 & Q==1) next
    if(P>=6 & Q==2) next
 

    cur.model <- 
        d.4stat%>% 
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * time.fct , random = ~1|flask, 
            correlation = corARMA(form = ~ Time | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}
pq.aic%>%arrange(aic)

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

I will test the best 5 models against AR1,

### compare AR1 with ARMA
```{r}

l.rm <- list()

pq.best <- pq.aic%>%
  slice_min( order_by=aic,n=5)

for (i in 1:nrow(pq.best)){
  l.rm[[i]] <- 
    d.4stat%>%
    # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * time.fct , 
        random = ~1|flask, 
        correlation = corARMA(form = ~ Time | flask,
                              p=pq.best$p[i],
                              q=pq.best$q[i]),
        data = .)
    
}


# add corAR1 to list
l.rm[[i+1]] <- 
  d.4stat%>%
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * time.fct ,
        random = ~1|flask, 
        correlation = corAR1(form = ~ Time | flask),
        data = .)


# get list order by AIC
aic.l <- sapply(l.rm,AIC)%>%
  sort.list()


# compare models by AIC order
anova.lme(l.rm[[aic.l[6]]],
          l.rm[[aic.l[5]]],
          l.rm[[aic.l[4]]],
          l.rm[[aic.l[3]]],
          l.rm[[aic.l[2]]],
          l.rm[[aic.l[1]]])
```

The best model, using corAR1, is significantly better than any of the other ARMA models.

```{r}
#display best model results
tmp <- summary(l.rm[[aic.l[6]]])
tmp$call
  anova.lme(l.rm[[aic.l[6]]])
```

The model coAR1 was chosen. 
**Phage and Time are significant as main effect and their interaction as well. Seed bank is mpt signifcant**


How sensetive would the result be to model selected?
```{r}
tmp <- map(l.rm, anova.lme)
tmp1 <- map(l.rm, getCall)

for(i in 1:6){
  print(tmp1[[i]])
  if(i<6) print(paste("p=",pq.best$p[i],"q=",pq.best$q[i]))
  
  print(tmp[[i]])
}
```

All models show the same story:  
* The interaction of interaction between phage and time is significant in influencing cell density. 


=================  

# Host density by host  

================= 

I will repeat the same anlysis but will treat the seed-bank as a 2-level factor: with (WT host) or without ($\Delta$spoIIE) a seed bank.

```{r}


# plot raw data
d.4stat%>%
    ggplot(aes(x=Time, y=cell.ml))+
      geom_line(aes(group=flask,color=host), size=1, alpha=0.7)+
      facet_grid(.~phage)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 1e10))

```

## Model selection  


### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r , message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the sappce in which the model convergence is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
    #this fllowing combination does not work: "iteration limit reached without convergence"
    if(P>=9 & Q==0) next
    if(P>=7 & Q==1) next


    cur.model <- 
        d.4stat%>% 
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * time.fct , random = ~1|flask, 
            correlation = corARMA(form = ~ Time | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}
pq.aic%>%arrange(aic)

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

I will test the best 5 models against AR1,

### compare AR1 with ARMA
```{r}

l.rm <- list()

pq.best <- pq.aic%>%
  slice_min( order_by=aic,n=5)

for (i in 1:nrow(pq.best)){
  l.rm[[i]] <- 
    d.4stat%>%
    # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * host * time.fct , random = ~1|flask, 
            correlation = corARMA(form = ~ Time | flask,
                                  p=pq.best$p[i],q=pq.best$q[i]),
            data = .)
    
}


# add corAR1 to list
l.rm[[i+1]] <- 
  d.4stat%>%
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * host * time.fct , random = ~1|flask, 
            correlation = corAR1(form = ~ Time | flask),
            data = .)


# get list order by AIC
aic.l <- sapply(l.rm,AIC)%>%
  sort.list()


# compare models by AIC order
anova.lme(l.rm[[aic.l[6]]],
          l.rm[[aic.l[5]]],
          l.rm[[aic.l[4]]],
          l.rm[[aic.l[3]]],
          l.rm[[aic.l[2]]],
          l.rm[[aic.l[1]]])
```


The best model, corAR1, is significantly better than any of the other ARMA models:

```{r}
#display best model results
getCall(l.rm[[aic.l[6]]])
  anova.lme(l.rm[[aic.l[6]]])
```

The model wirth coAR1 was chosen. 
**Phage and Time are significant as main effect and their interaction as well. Seed bank is mpt signifcant**


How sensetive would the result be to model selected?
```{r}
tmp <- map(l.rm, anova.lme)
tmp1 <- map(l.rm, getCall)

for(i in 1:6){
  print(tmp1[[i]])
  if(i<6) print(paste("p=",pq.best$p[i],"q=",pq.best$q[i]))
  
  print(tmp[[i]])
}
```

All models show the same story: as sbove.


=================  

# phage density  

=================   

```{r}
d.4stat%>%
  filter(phage=="SPO1")%>%
    ggplot(aes(x=Time, y=phage.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      theme_bw()+
      panel_border()+
      # scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("phage/ml (log)")+
      coord_cartesian(ylim = c(7, 10))
```

Here I will remove time 0, the input phage

```{r}
d.4stat%>%
  filter(phage=="SPO1")%>%
  filter(Time!=0)%>%
    ggplot(aes(x=Time, y=phage.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("phage/ml (log)")+
      coord_cartesian(ylim = c(5e7, 1e10))
```

## Model selection  

### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r , message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the space in which the model convergence is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
        #this fllowing combination does not work: "iteration limit reached without convergence"
    if(P>=10 & Q==0) next
    if(P==4 & Q==1) next
    if(P>=8 & Q==1) next
    
    cur.model <- 
        d.4stat%>% 
      filter(Time!=0)%>%
      filter(phage=="SPO1")%>%
  # looking only at total cells (cell.ml)
    lme(phage.ml ~ seed.bank * time.fct , random = ~1|flask, 
            correlation = corARMA(form = ~ Time | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}

pq.aic%>%arrange(aic)

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

I will test the best 5 models against AR1,

### compare AR1 with ARMA
```{r}

l.rm <- list()

pq.best <- pq.aic%>%
  slice_min( order_by=aic,n=5)

for (i in 1:nrow(pq.best)){
  l.rm[[i]] <- 
    d.4stat%>% 
    filter(Time!=0)%>%
    filter(phage=="SPO1")%>%
  # looking only at phages (phage.ml)
    lme(phage.ml ~ seed.bank * time.fct , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask,
                                  p=pq.best$p[i],q=pq.best$q[i]),
            data = .)
    
}


# add corAR1 to list
l.rm[[i+1]] <- 
  d.4stat%>%
  filter(phage=="SPO1")%>%
  filter(Time!=0)%>%
 # looking only at phages (phage.ml)
    lme(phage.ml ~ seed.bank * time.fct , random = ~1|flask, 
            correlation = corAR1(form = ~ 1 | flask),
            data = .)


# get list order by AIC
aic.l <- sapply(l.rm,AIC)%>%
  sort.list()


# compare models by AIC order
anova.lme(l.rm[[aic.l[6]]],
          l.rm[[aic.l[5]]],
          l.rm[[aic.l[4]]],
          l.rm[[aic.l[3]]],
          l.rm[[aic.l[2]]],
          l.rm[[aic.l[1]]])
```

Here too, the coAR1 is signicantly better.

```{r}
#display best model results
getCall(l.rm[[aic.l[6]]])
  anova(l.rm[[aic.l[6]]])
```

**There is a significant effect of seed.bank and time on phage population density.**

How sensetive would the result be to model selected?
```{r}
tmp <- map(l.rm, anova.lme)
tmp1 <- map(l.rm, getCall)

for(i in 1:6){
  print(tmp1[[i]])
  if(i<6) print(paste("p=",pq.best$p[i],"q=",pq.best$q[i]))
  
  print(tmp[[i]])
}
```

All models show the same story:  Seed bank X time significant for phage.

=================  

# spore density  

=================  

```{r}
d.4stat%>%
  filter(host=="WT")%>%
       ggplot(aes(x=Time, y=spore.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
  geom_hline(yintercept = detection.limit, color="red")+
  geom_label(label="detection limit", x=10, y=log10(detection.limit), color="red")+
      facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("spore/ml (log)")+
      coord_cartesian(ylim = c(1e5, 2e9))
```

All points usable.

## Model selection  

### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r , message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the space in which the model convergence is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
        #this fllowing combination does not work: "iteration limit reached without convergence"
    if(P>=6 & Q==2) next
    
    cur.model <- 
        d.4stat%>% 
      filter(host=="WT")%>%
  # looking only at total cells (cell.ml)
    lme(spore.ml ~ phage*seed.bank * time.fct , random = ~1|flask, 
            correlation = corARMA(form = ~ Time | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}

pq.aic%>%arrange(aic)

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

I will test the best 5 models against AR1,

### compare AR1 with ARMA
```{r}

l.rm <- list()

pq.best <- pq.aic%>%
  slice_min( order_by=aic,n=5)

for (i in 1:nrow(pq.best)){
  l.rm[[i]] <- 
        d.4stat%>% 
      filter(host=="WT")%>%
  # looking only at total spore (spore.ml)
    lme(spore.ml ~ phage*seed.bank * time.fct , random = ~1|flask,
        correlation = corARMA(form = ~ 1 | flask,
                              p=pq.best$p[i],q=pq.best$q[i]),
            data = .)
    
}


# add corAR1 to list
l.rm[[i+1]] <- 
        d.4stat%>% 
      filter(host=="WT")%>%
  # looking only at total spore (spore.ml)
    lme(spore.ml ~ phage*seed.bank * time.fct , random = ~1|flask,
            correlation = corAR1(form = ~ 1 | flask),
            data = .)


# get list order by AIC
aic.l <- sapply(l.rm,AIC)%>%
  sort.list()


# compare models by AIC order
anova.lme(l.rm[[aic.l[6]]],
          l.rm[[aic.l[5]]],
          l.rm[[aic.l[4]]],
          l.rm[[aic.l[3]]],
          l.rm[[aic.l[2]]],
          l.rm[[aic.l[1]]])
```

Once agin the corAR1 model is best.


```{r}
#display best model results
getCall(l.rm[[aic.l[6]]])
  anova.lme(l.rm[[aic.l[6]]])
```

**There is significant effect of  interaction of phage, seed.bank and time  on spore population density.**

How sensetive would the result be to model selected?
```{r}
tmp <- map(l.rm, anova.lme)
tmp1 <- map(l.rm, getCall)

for(i in 1:6){
  print(tmp1[[i]])
  if(i<6) print(paste("p=",pq.best$p[i],"q=",pq.best$q[i]))
  
  print(tmp[[i]])
}
```

The phage:seed-bank:time interaction is hovering closely below 0.05, but in all case ur us below it.



=================  

# percent spore  

=================  

```{r}
d.4stat <- d.4stat%>%
  mutate(perc.spore=100*spore.ml/cell.ml)

d.4stat%>%
  filter(host=="WT")%>%
    ggplot(aes(x=Time, y=perc.spore))+
    geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
    theme_bw()+
    panel_border()+
  facet_wrap(~phage)+
  theme(legend.position = "bottom",
            text=element_text(size=14))+
  ylab("% spore")+
  ylim(0,100)
```

All points usable.

## Model selection  

### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r , message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the space in which the model convergence is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
        #this fllowing combination does not work: "iteration limit reached without convergence"
    if(P>=10 & Q==0) next
    if(P>=6 & Q==1) next
    if(P>=6 & Q==2) next
    
    cur.model <- 
        d.4stat%>% 
      filter(host=="WT")%>%
  # looking only at total cells (cell.ml)
    lme(perc.spore ~ phage*seed.bank * time.fct , random = ~1|flask, 
            correlation = corARMA(form = ~ Time | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}

pq.aic%>%arrange(aic)

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

I will test the best 5 models against AR1,

### compare AR1 with ARMA
```{r}
 
l.rm <- list()

pq.best <- pq.aic%>%
  slice_min( order_by=aic,n=5)

for (i in 1:nrow(pq.best)){
  l.rm[[i]] <- 
        d.4stat%>% 
      filter(host=="WT")%>%
  # looking only at total spore (spore.ml)
    lme(perc.spore ~ phage*seed.bank * time.fct , random = ~1|flask,
        correlation = corARMA(form = ~ Time | flask,
                              p=pq.best$p[i],q=pq.best$q[i]),
            data = .)
    
}


# add corAR1 to list
l.rm[[i+1]] <- 
        d.4stat%>% 
      filter(host=="WT")%>%
  # looking only at total spore (spore.ml)
    lme(perc.spore ~ phage*seed.bank * time.fct , random = ~1|flask,
            correlation = corAR1(form = ~ Time | flask),
            data = .)


# get list order by AIC
aic.l <- sapply(l.rm,AIC)%>%
  sort.list()


# compare models by AIC order
anova.lme(l.rm[[aic.l[6]]],
          l.rm[[aic.l[5]]],
          l.rm[[aic.l[4]]],
          l.rm[[aic.l[3]]],
          l.rm[[aic.l[2]]],
          l.rm[[aic.l[1]]])
```

The first model, corAR1, is significantly better than the second one. 

```{r}
#display best model results
getCall(l.rm[[aic.l[6]]])

  anova.lme(l.rm[[aic.l[6]]])
```

**There is a significant interaction of  2-way interactions with time on percent spore in the population, but not for the 3-way or for phageXseed bank.**

How sensetive would the result be to model selected?
```{r}
tmp <- map(l.rm, anova.lme)
tmp1 <- map(l.rm, getCall)

for(i in 1:6){
  print(tmp1[[i]])
  if(i<6) print(paste("p=",pq.best$p[i],"q=",pq.best$q[i]))
  
  print(tmp[[i]])
}
```

A coiple of the models (no. 2 and 3) hav the 2-way intercaction of phage X seed bank significant, and all models have the 3-way interaction as non significant.

=================  

# Veg cells  

=================  

```{r}
d.4stat%>%
    ggplot(aes(x=Time, y=veg.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
  geom_hline(yintercept = detection.limit, color="red")+
  geom_label(label="detection limit", x=10, y=log10(detection.limit), color="red")+
      facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("veg. cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 2e9))
```

All points usable.

## Model selection  

### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r , message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the space in which the model convergence is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
        #this fllowing combination does not work: "iteration limit reached without convergence"
    if(P>=9 & Q==0) next
    if(P>=8 & Q==1) next
    # if(P==1 & Q==2) next
    # if(P==3 & Q==2) next
    
    cur.model <- 
        d.4stat%>% 
  # looking only at veg cells (veg.ml)
    lme(veg.ml ~ phage*seed.bank * time.fct , random = ~1|flask, 
            correlation = corARMA(form = ~ Time | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}

pq.aic%>%arrange(aic)

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

I will test the best 5 models against AR1,

### compare AR1 with ARMA
```{r}
 
l.rm <- list()

pq.best <- pq.aic%>%
  slice_min( order_by=aic,n=5)

for (i in 1:nrow(pq.best)){
  l.rm[[i]] <- 
        d.4stat%>% 
  # looking only at veg cells (veg.ml)
    lme(veg.ml ~ phage*seed.bank * time.fct , random = ~1|flask,
        correlation = corARMA(form = ~ Time | flask,
                              p=pq.best$p[i],q=pq.best$q[i]),
            data = .)
    
}


# add corAR1 to list
l.rm[[i+1]] <- 
        d.4stat%>% 
  # looking only at veg cells (veg.ml)
    lme(veg.ml ~ phage*seed.bank * time.fct , random = ~1|flask,
            correlation = corAR1(form = ~ Time | flask),
            data = .)


# get list order by AIC
aic.l <- sapply(l.rm,AIC)%>%
  sort.list()


# compare models by AIC order
anova.lme(l.rm[[aic.l[6]]],
          l.rm[[aic.l[5]]],
          l.rm[[aic.l[4]]],
          l.rm[[aic.l[3]]],
          l.rm[[aic.l[2]]],
          l.rm[[aic.l[1]]])
```

The first model (no. 6), corAR1, is significantly better than the second one. 

```{r}
#display best model results
getCall(l.rm[[aic.l[6]]])

  anova.lme(l.rm[[aic.l[6]]])
```

**There is significant effect of interaction of phage X seed.bank and phage X time on veg population density, but not 3-way.**

How sensetive would the result be to model selected?
```{r}
tmp <- map(l.rm, anova.lme)
tmp1 <- map(l.rm, getCall)

for(i in 1:6){
  print(tmp1[[i]])
  if(i<6) print(paste("p=",pq.best$p[i],"q=",pq.best$q[i]))
  
  print(tmp[[i]])
}
```

The other models also have those interactions, ut some also show the 3-way to be significant.