---
title: "RM-ANOVA for coevolution with a seed bank"
output: html_notebook
editor_options: 
  chunk_output_type: inline
  code_folding: hide
---

```{r setup, include=FALSE}
library(renv)
# install to project library the required packages
renv::restore()
# # Initialize lock file when first setting repo
# renv::init()

library(here)
# # set root at project file ("coevolution-ts"). 
# # Done once at initial setup.
# here::set_here()

library(tidyverse)
library(cowplot)


library(nlme)

# save the state of the project library to the lockfile (called renv.lock),
renv::snapshot()

```

## Population density data  
I have previously consolidated all the data on population density collected for this experiment. This includes:

| Population | What it is | Method of measurement |
| - | - | - |
|spore / mL | host spore density | flow-cytometry: low SYBR population |
|veg / mL | host vegetative cell density | flow-cytometry: high SYBR population |
| cell /mL | total host cell density (spore+veg) | flow-cytometry
| phage/ mL | total SPO1 phage density (free phage) | qPCR (SYBR)  with phage specific primers and lysate serial dilution as standard | 

```{r load data, echo=FALSE, message=FALSE}
d <- read_csv(here("data","coevolution_pop_density.csv"), )

#add columns on host strain and phage treatment  as encoded in culture name
d <- 
  d%>%
  mutate(host=case_when(grepl("W", culture) ~ "WT",
                         grepl("SN", culture) ~ "dSpoIIE",
                         grepl("dS", culture) ~ "dSpoIIE"))%>%
  
  mutate(phage=case_when(grepl("O", culture) ~ "SPO1",
                         grepl("Ct", culture) ~ "no_Phage",
                         grepl("ANC", line) ~ "no_Phage"))%>%

   mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))%>%
  #add flask to be used as grouping variable
  mutate(flask=paste(culture,line, sep = "-"))%>%
  #remove setup data
  filter(Time>=0)

# plot raw data
d%>%
    pivot_longer(cols=contains(".ml"), names_to="pop",values_to="num.ml")%>%
    ggplot(aes(x=Time, y=num.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      facet_grid(pop~phage)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 1e10))
```

In the first few days of the experiment cell numbers dropped below detection limit in 2 of 3 SNO lines. Missing data cannot be analyzed in RM-ANOVA. I will assign these points the detection limit values.

Points with missing data: 
```{r}
d%>%
  filter(cell.ml<10)
```


---  

**Flow-cytometry detection limit**

The flow cytometer stop conditions were either $50\mu L$ or 50,000 events, whatever came first.  
In addition, to account for false asignments by noise I required that a gated population have at least 100 events to be taken into account. So the minimal detected population would have 100 cells in $50\mu L$ of analyzed liquid. At this poimt of the experiment we were analyzing x100 diluted samples.

```{r}
detection.limit <- 100 * (1000/50) * 100
```

100 events X $1000/50\mu L$ X100 = `r detection.limit` 

I will use $1X10^5$, half the detection limit dor missing value points
---




```{r}
d <- d%>%
  mutate(cell.ml=if_else(cell.ml>10, cell.ml,detection.limit/2))
# plot total cells of raw data
d%>%
    ggplot(aes(x=Time, y=cell.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
  geom_hline(yintercept = detection.limit, color="red")+
  geom_label(label="detection limit", x=10, y=log10(detection.limit), color="red")+
      facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 2e9))
```

```{r}

# plot total cells of raw data
d%>%
  group_by(phage,seed.bank,Time)%>%
  summarise(n=n(),sd=sd(cell.ml), cell.ml=mean(cell.ml), sem=sd/sqrt(n))%>%
    ggplot(aes(x=Time, y=cell.ml))+
  geom_errorbar(aes(ymin=cell.ml-sem,ymax=cell.ml+sem,color=seed.bank), alpha=0.7)+
      geom_line(aes(color=seed.bank), size=1, alpha=0.7)+
      facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 2e9))
```

<!-- # Auto correlation -->
<!-- Based on Zuur, Alain F., et al. "Violation of Independence–Part I." Mixed effects models and extensions in ecology with R. Springer, New York, NY, 2009. 143-160. -->
<!-- ```{r} -->
<!-- #no auto correlation model -->
<!-- M0 <-     lme(cell.ml ~seed.bank*Time, random = ~1|flask , -->
<!--                      data = d%>%filter(phage=="SPO1")) -->

<!-- E <- residuals(M0, type = "normalized") -->
<!-- acf(E, main="SPO1")  -->

<!-- #no auto correation model -->
<!-- M0 <-     lme(cell.ml ~seed.bank*Time, random = ~1|flask , -->
<!--                      data = d%>%filter(phage!="SPO1")) -->

<!-- E <- residuals(M0, type = "normalized") -->
<!-- acf(E, main="no Phage")  -->


<!-- ``` -->

# Adjusting variable types for lme

In the lme models I will be using the fixd effects need to be secified as factore, in this experiment these ar: *Phage, seed.bank and time*. However for the autocorrelation specification time needs to be specified as an integer. To fulfill both requirements I will use for the time unit the experimental day rather than transfer, since we measured onc a day but twice per transfer. This will be simplty *time x 2*. I will then make fom that a separate varible which will be the factor of the time.

```{r}
d.4stat <- 
  d%>%
  # remove setup data
  filter(Time>=0)%>%
  #make time an integer for correlation structure
  # by converting to days as time unit
  mutate(Time=as.integer(2*Time))%>%
  # for the lme model all fixed effects need to be factors
  mutate(time.fct=as.factor(Time))%>%
  mutate(seed.bank=as.factor(seed.bank))%>%
  mutate(phage=as.factor(phage))
```


## Model selection  


### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r arma, message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the space in which the model conerges is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
    #this fllowing combination does not work: "iteration limit reached without convergence"
    if(P>=8 & Q==0) next
     if(P>=7 & Q==1) next
    if(P==1 & Q==2) next
    if(P>=6 & Q==2) next

    cur.model <- 
        d.4stat%>% 
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * time.fct , random = ~1|flask, 
            correlation = corARMA(form = ~ Time | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}
pq.aic%>%arrange(aic)

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

I will test the best 5 models against AR1,

### compare AR1 with ARMA
```{r}

l.rm <- list()

pq.best <- pq.aic%>%
  slice_min( order_by=aic,n=5)

for (i in 1:nrow(pq.best)){
  l.rm[[i]] <- 
    d.4stat%>%
    # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * time.fct , 
        random = ~1|flask, 
        correlation = corARMA(form = ~ Time | flask,
                              p=pq.best$p[i],
                              q=pq.best$q[i]),
        data = .)
    
}


# add corAR1 to list
l.rm[[i+1]] <- 
  d.4stat%>%
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * time.fct ,
        random = ~1|flask, 
        correlation = corAR1(form = ~ Time | flask),
        data = .)


# get list order by AIC
aic.l <- sapply(l.rm,AIC)%>%
  sort.list()


# compare models by AIC order
anova.lme(l.rm[[aic.l[6]]],
          l.rm[[aic.l[5]]],
          l.rm[[aic.l[4]]],
          l.rm[[aic.l[3]]],
          l.rm[[aic.l[2]]],
          l.rm[[aic.l[1]]])
```

The best model is not significantly better than any of the other ARMA models. Looking at model 6 (corAR1), has the least degrees of freedom (df=177) :

```{r}
#display best model results
tmp <- summary(l.rm[[aic.l[6]]])
tmp$call
  car::Anova(l.rm[[aic.l[6]]])
```

The model coAR1 was chosen. 
**There is a significant interaction between phage,seed.bank and time effecting population density.**


How sensetive would the result be to model selected?
```{r}
tmp <- map(l.rm, car::Anova)
tmp1 <- map(l.rm, getCall)

for(i in 1:6){
  print(tmp1[[i]])
  if(i<6) print(paste("p=",pq.best$p[i],"q=",pq.best$q[i]))
  
  print(tmp[[i]])

  print("--+--+--+--+--+--+--+--")
  print("                       ")
}
```

All models show the same story:  
* The interaction of interaction between phage,seed.bank and time is significant in influencing cell density. The selected model (#6) actually has the highest (i.e. "worst") p-value for the 3 way interaction.

<!-- ## Post hoc -->


<!-- comparing means across seed bank:  -->
<!-- ```{r, warning=F, message=F} -->
<!-- source(here("code","functions","plot_TukeyHSD.R")) -->

<!-- tuk.p <- TukeyHSD(aov(cell.ml~seed.bank, -->
<!--                     data =d.4stat%>% -->
<!--                       filter(phage=="SPO1")))%>% -->
<!--   plot_TukeyHSD(back = "return") -->


<!-- tuk.nop <- TukeyHSD(aov(cell.ml~seed.bank, -->
<!--                     data =d.4stat%>% -->
<!--                       filter(phage!="SPO1")))%>% -->
<!--   plot_TukeyHSD(back = "return") -->

<!-- plot_grid(tuk.p,tuk.nop,labels = c("SPO1","no Phage"), nrow = 2) -->

<!-- # TukeyHSD(aov(cell.ml~seed.bank*phage*time.fct,data =d.4stat))%>% -->
<!-- #                       plot_TukeyHSD(back = "plot") -->

<!-- ``` -->


=================  

# Host density by host  

================= 

I will repeat the same anlysis but will treat the seed-bank as a 2-level factor: with (WT host) or without ($\Delta$spoIIE) a seed bank.

## Model selection  


### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r , message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the sappce in which the model convergence is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-3*Q))){ 
    
    #this fllowing combination does not work: "iteration limit reached without convergence"
    if(P>=1 & Q==2) next
    if(P>=8 & Q==0) next
    if(P>=7 & Q==1) next
    # if(P>=3 & Q==2) next

    cur.model <- 
        d.4stat%>% 
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * time.fct , random = ~1|flask, 
            correlation = corARMA(form = ~ Time | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}
pq.aic%>%arrange(aic)

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

I will test the best 5 models against AR1,

### compare AR1 with ARMA
```{r}

l.rm <- list()

pq.best <- pq.aic%>%
  slice_min( order_by=aic,n=5)

for (i in 1:nrow(pq.best)){
  l.rm[[i]] <- 
    d.4stat%>%
    # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * host * time.fct , random = ~1|flask, 
            correlation = corARMA(form = ~ Time | flask,
                                  p=pq.best$p[i],q=pq.best$q[i]),
            data = .)
    
}


# add corAR1 to list
l.rm[[i+1]] <- 
  d.4stat%>%
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * host * time.fct , random = ~1|flask, 
            correlation = corAR1(form = ~ Time | flask),
            data = .)


# get list order by AIC
aic.l <- sapply(l.rm,AIC)%>%
  sort.list()


# compare models by AIC order
anova.lme(l.rm[[aic.l[6]]],
          l.rm[[aic.l[5]]],
          l.rm[[aic.l[4]]],
          l.rm[[aic.l[3]]],
          l.rm[[aic.l[2]]],
          l.rm[[aic.l[1]]])
```

The best model is not significantly better than any of the other ARMA models. Looking at model 4, has the least degrees of freedom (df=119), This is the coAR1 model :

```{r}
#display best model results
getCall(l.rm[[aic.l[4]]])
  car::Anova(l.rm[[aic.l[4]]])
```

The model wirth coAR1 was chosen. 
**There is a significant interaction between phage,seed.bank and time effecting population density.**


How sensetive would the result be to model selected?
```{r}
tmp <- map(l.rm, car::Anova)
tmp1 <- map(l.rm, getCall)

for(i in 1:6){
  print(tmp1[[i]])
  if(i<6) print(paste("p=",pq.best$p[i],"q=",pq.best$q[i]))
  
  print(tmp[[i]])

  print("--+--+--+--+--+--+--+--")
  print("                       ")
}
```

All models show the same story: as sbove.


=================  

# phage density  

=================   

```{r}
d%>%
  filter(phage=="SPO1")%>%
    ggplot(aes(x=Time, y=phage.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("phage/ml (log)")+
      coord_cartesian(ylim = c(1e5, 1e10))
```

Here I will remove time 0, the input phage

```{r}
d%>%
  filter(phage=="SPO1")%>%
  filter(Time!=0)%>%
    ggplot(aes(x=Time, y=phage.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("phage/ml (log)")+
      coord_cartesian(ylim = c(1e5, 1e10))
```

## Model selection  

### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r , message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the space in which the model convergence is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
        #this fllowing combination does not work: "iteration limit reached without convergence"
    if(P==5 & Q==2) next
    
    cur.model <- 
        d%>% 
      filter(phage=="SPO1")%>%
  # looking only at total cells (cell.ml)
    lme(phage.ml ~ seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}

pq.aic%>%arrange(aic)

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

I will test the best 5 models against AR1,

### compare AR1 with ARMA
```{r}

l.rm <- list()

pq.best <- pq.aic%>%
  slice_min( order_by=aic,n=5)

for (i in 1:nrow(pq.best)){
  l.rm[[i]] <- 
    d%>%  
    filter(phage=="SPO1")%>%
  # looking only at phages (phage.ml)
    lme(phage.ml ~ seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask,
                                  p=pq.best$p[i],q=pq.best$q[i]),
            data = .)
    
}


# add corAR1 to list
l.rm[[i+1]] <- 
  d%>%
  filter(phage=="SPO1")%>%
 # looking only at phages (phage.ml)
    lme(phage.ml ~ seed.bank * Time , random = ~1|flask, 
            correlation = corAR1(form = ~ 1 | flask),
            data = .)


# get list order by AIC
aic.l <- sapply(l.rm,AIC)%>%
  sort.list()


# compare models by AIC order
anova.lme(l.rm[[aic.l[6]]],
          l.rm[[aic.l[5]]],
          l.rm[[aic.l[4]]],
          l.rm[[aic.l[3]]],
          l.rm[[aic.l[2]]],
          l.rm[[aic.l[1]]])
```

The first 2 best models are comparable but significantly better than the third one. Looking at model with lower DF, model 1:

```{r}
#display best model results
summary(l.rm[[aic.l[1]]])
  anova.lme(l.rm[[aic.l[1]]])
```

**No significant effect of seed.bank on phage population density at any level.**

How sensetive would the result be to model selected?
```{r}
map(l.rm,anova.lme)
```

All models show the same story:  Seed bank is not significant for phage.

=================  

# spore density  

=================  

```{r}
d%>%
  filter(host=="WT")%>%
    ggplot(aes(x=Time, y=spore.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
  facet_wrap(~phage)+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("spore/ml (log)")+
      coord_cartesian(ylim = c(1e5, 1e10))
```

All points usable.

## Model selection  

### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r , message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the space in which the model convergence is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
        #this fllowing combination does not work: "iteration limit reached without convergence"
    if(P>=5 & Q==2) next
    
    cur.model <- 
        d%>% 
      filter(host=="WT")%>%
  # looking only at total cells (cell.ml)
    lme(spore.ml ~ phage*seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}

pq.aic%>%arrange(aic)

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

I will test the best 5 models against AR1,

### compare AR1 with ARMA
```{r}

l.rm <- list()

pq.best <- pq.aic%>%
  slice_min( order_by=aic,n=5)

for (i in 1:nrow(pq.best)){
  l.rm[[i]] <- 
        d%>% 
      filter(host=="WT")%>%
  # looking only at total spore (spore.ml)
    lme(spore.ml ~ phage*seed.bank * Time , random = ~1|flask,
        correlation = corARMA(form = ~ 1 | flask,
                              p=pq.best$p[i],q=pq.best$q[i]),
            data = .)
    
}


# add corAR1 to list
l.rm[[i+1]] <- 
        d%>% 
      filter(host=="WT")%>%
  # looking only at total spore (spore.ml)
    lme(spore.ml ~ phage*seed.bank * Time , random = ~1|flask,
            correlation = corAR1(form = ~ 1 | flask),
            data = .)


# get list order by AIC
aic.l <- sapply(l.rm,AIC)%>%
  sort.list()


# compare models by AIC order
anova.lme(l.rm[[aic.l[6]]],
          l.rm[[aic.l[5]]],
          l.rm[[aic.l[4]]],
          l.rm[[aic.l[3]]],
          l.rm[[aic.l[2]]],
          l.rm[[aic.l[1]]])
```

The first 2 best models are comparable and not significantly better than the third one. Looking at model with lower DF, model 1:

```{r}
#display best model results
summary(l.rm[[aic.l[1]]])
  anova.lme(l.rm[[aic.l[1]]])
```

**No significant effect of seed.bank treatment on spore population density. However at a 0.1 significancIIe level, there is a seed-bank intraction with time.**

How sensetive would the result be to model selected?
```{r}
map(l.rm,anova.lme)
```

The seed-bank:time interaction is hovering around 0.05, some models have it above and some below.


## Post hoc

comparing means across seed bank: 
```{r, warning=F, message=F}
source(here("code","functions","plot_TukeyHSD.R"))

tuk.p <- TukeyHSD(aov(spore.ml~seed.bank,
                    data =d%>%
                      filter(host=="WT")%>%
                      filter(phage=="SPO1")))%>%
  plot_TukeyHSD(back = "return")


tuk.nop <- TukeyHSD(aov(spore.ml~seed.bank,
                    data =d%>%
                      filter(host=="WT")%>%
                      filter(phage!="SPO1")))%>%
  plot_TukeyHSD(back = "return")

plot_grid(tuk.p,tuk.nop,labels = c("SPO1","no Phage"), nrow = 2)
```

In the no-phage lines the confidence interval for the difference between seed-bank treatments is just touching 0. This is likely the sourc of the borderline significance osbserved.


=================  

# percent spore  

=================  

```{r}
d <- d%>%
  mutate(perc.spore=100*spore.ml/cell.ml)

d%>%
  filter(host=="WT")%>%
    ggplot(aes(x=Time, y=perc.spore))+
    geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
    theme_bw()+
    panel_border()+
  facet_wrap(~phage)+
  theme(legend.position = "bottom",
            text=element_text(size=14))+
  ylab("% spore")+
  ylim(0,100)
```

All points usable.

## Model selection  

### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r , message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the space in which the model convergence is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
        #this fllowing combination does not work: "iteration limit reached without convergence"
    if(P>=4 & Q==2) next
    
    cur.model <- 
        d%>% 
      filter(host=="WT")%>%
  # looking only at total cells (cell.ml)
    lme(perc.spore ~ phage*seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}

pq.aic%>%arrange(aic)

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

I will test the best 5 models against AR1,

### compare AR1 with ARMA
```{r}

l.rm <- list()

pq.best <- pq.aic%>%
  slice_min( order_by=aic,n=5)

for (i in 1:nrow(pq.best)){
  l.rm[[i]] <- 
        d%>% 
      filter(host=="WT")%>%
  # looking only at total spore (spore.ml)
    lme(perc.spore ~ phage*seed.bank * Time , random = ~1|flask,
        correlation = corARMA(form = ~ 1 | flask,
                              p=pq.best$p[i],q=pq.best$q[i]),
            data = .)
    
}


# add corAR1 to list
l.rm[[i+1]] <- 
        d%>% 
      filter(host=="WT")%>%
  # looking only at total spore (spore.ml)
    lme(perc.spore ~ phage*seed.bank * Time , random = ~1|flask,
            correlation = corAR1(form = ~ 1 | flask),
            data = .)


# get list order by AIC
aic.l <- sapply(l.rm,AIC)%>%
  sort.list()


# compare models by AIC order
anova.lme(l.rm[[aic.l[6]]],
          l.rm[[aic.l[5]]],
          l.rm[[aic.l[4]]],
          l.rm[[aic.l[3]]],
          l.rm[[aic.l[2]]],
          l.rm[[aic.l[1]]])
```

The first 2 best models are comparable and not significantly better than the third one. Looking at model with lower DF, model 1:

```{r}
#display best model results
summary(l.rm[[aic.l[1]]])
  anova.lme(l.rm[[aic.l[1]]])
```

**No significant effect of seed.bank treatment on percent spore in the population density.**

How sensetive would the result be to model selected?
```{r}
map(l.rm,anova.lme)
```

All models have no significant effects for seed bank on percent spores.