---
title: "RM-ANOVA for coevolution with a seed bank"
output: html_notebook
editor_options: 
  chunk_output_type: inline
  code_folding: hide
---

```{r setup, include=FALSE}
library(renv)
# install to project library the required packages
renv::restore()
# # Initialize lock file when first setting repo
# renv::init()

library(here)
# # set root at project file ("coevolution-ts"). 
# # Done once at initial setup.
# here::set_here()

library(tidyverse)
library(cowplot)


library(nlme)

# save the state of the project library to the lockfile (called renv.lock),
renv::snapshot()

```

## Population density data  
I have previously consolidated all the data on population density collected for this experiment. This includes:

| Population | What it is | Method of measurement |
| - | - | - |
|spore / mL | host spore density | flow-cytometry: low SYBR population |
|veg / mL | host vegetative cell density | flow-cytometry: high SYBR population |
| cell /mL | total host cell density (spore+veg) | flow-cytometry
| phage/ mL | total SPO1 phage density (free phage) | qPCR (SYBR)  with phage specific primers and lysate serial dilution as standard | 

```{r load data, echo=FALSE, message=FALSE}
d <- read_csv(here("data","coevolution_pop_density.csv"), )

#add columns on host strain and phage treatment  as encoded in culture name
d <- 
  d%>%
  mutate(host=case_when(grepl("W", culture) ~ "WT",
                         grepl("SN", culture) ~ "dSpoIIE",
                         grepl("dS", culture) ~ "dSpoIIE"))%>%
  
  mutate(phage=case_when(grepl("O", culture) ~ "SPO1",
                         grepl("Ct", culture) ~ "no_Phage",
                         grepl("ANC", line) ~ "no_Phage"))%>%

   mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))%>%
  #add flask to be used as grouping variable
  mutate(flask=paste(culture,line, sep = "-"))%>%
  #remove setup data
  filter(Time>=0)

# plot raw data
d%>%
    pivot_longer(cols=contains(".ml"), names_to="pop",values_to="num.ml")%>%
    ggplot(aes(x=Time, y=num.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      facet_grid(pop~phage)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 1e10))
```

In the first few days of the experiment cell numbers dropped below detection limit in 2 of 3 SNO lines. Missing data cannot be analyzed in RM-ANOVA. I will assign these points the detection limit values.

Points with missing data: 
```{r}
d%>%
  filter(cell.ml<10)
```
 
 These are all veg cells:

```{r}
d%>%
  filter(veg.ml<10)
```
---  

**Flow-cytometry detection limit**

The flow cytometer stop conditions were either $50\mu L$ or 50,000 events, whatever came first.  
In addition, to account for false asignments by noise I required that a gated population have at least 100 events to be taken into account. So the minimal detected population would have 100 cells in $50\mu L$ of analyzed liquid. At this poimt of the experiment we were analyzing x100 diluted samples.

```{r}
detection.limit <- 100 * (1000/50) * 100
```

100 events X $1000/50\mu L$ X100 = `r detection.limit` 

I will use $1X10^5$, half the detection limit dor missing value points
---

```{r}
d <- d%>%
  mutate(cell.ml=if_else(cell.ml>10, cell.ml,detection.limit/2))%>%
  mutate(veg.ml=if_else(veg.ml>10, veg.ml,detection.limit/2))


# plot total cells of raw data
d%>%
    ggplot(aes(x=Time, y=cell.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
  geom_hline(yintercept = detection.limit, color="red")+
  geom_label(label="detection limit", x=10, y=log10(detection.limit), color="red")+
      facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 2e9))
```


# Adjusting variable types for lme

In the lme models I will be using the fixd effects need to be secified as factore, in this experiment these ar: *Phage, seed.bank and time*. However for the autocorrelation specification time needs to be specified as an integer. To fulfill both requirements I will use for the time unit the experimental day rather than transfer, since we measured onc a day but twice per transfer. This will be simplty *time x 2*. I will then make fom that a separate varible which will be the factor of the time.

```{r}
d.4stat <- 
  d%>%
  # remove setup data
  filter(Time>=0)%>%
  #make time an integer for correlation structure
  # by converting to days as time unit
  mutate(Time=as.integer(2*Time))%>%
  # for the lme model all fixed effects need to be factors
  mutate(time.fct=as.factor(Time))%>%
  mutate(seed.bank=as.factor(seed.bank))%>%
  mutate(phage=as.factor(phage))
```

# test data for homogeneity of variances
```{r}
# Levene's test with multiple independent variables
car::leveneTest(cell.ml ~ phage*seed.bank*line, data = d.4stat)
```


# RM-Anova with  corAR1 covariance structure

```{r}
rm.ar1 <- 
  d.4stat%>%
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * time.fct ,
    # lme(log10(cell.ml) ~ phage * seed.bank * time.fct ,
        random = ~1|flask, 
        correlation = corAR1(form = ~ Time | flask),
        data = .)

anova.lme(rm.ar1)
```


Looking at residuals  
```{r}
broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(Time,.resid)) +
  geom_hline(yintercept = 0)+
  geom_line(aes(color=line))+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(seed.bank,.resid)) +
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(fill=line),alpha=.5, position = position_dodge(width = .9))+
  geom_point(aes(color=line), position = position_dodge(width = .9), shape=21)+
  facet_wrap(~phage)+
  theme_cowplot()+panel_border()

broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(x = .resid, color = line))+
  geom_density()+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

```

```{r}
d.resid <- 
  broom.mixed::augment(rm.ar1)%>%
  select(.resid,phage,seed.bank,line)%>%
  group_by(phage,seed.bank,line)%>%
summarise(p.shapiro = shapiro.test(unlist(.resid))$p.value)%>%
  mutate(normal=if_else(p.shapiro>=0.05,TRUE,FALSE))%>%
  arrange(desc(p.shapiro))%>%
  print()

n.norm=sum(d.resid$normal)

broom.mixed::augment(rm.ar1)%>%
  select(.resid,phage,seed.bank,line)%>%
  group_by(phage,seed.bank,line)%>%
summarise(p.shapiro = shapiro.test(unlist(.resid))$p.value)%>%
  mutate(normal=if_else(p.shapiro>=0.05,TRUE,FALSE))%>%
  arrange(desc(p.shapiro))%>%
  ungroup()%>%
  mutate(rank=row_number())%>%
  ggplot(aes(rank, p.shapiro))+
  geom_col(aes(fill=seed.bank, color=phage),size=1)+
  geom_hline(yintercept = 0.05, color="red")+
  geom_vline(xintercept = n.norm+.5, color="blue")+
  scale_fill_grey()+
  scale_y_log10()+
  theme_cowplot()
```


# log transformed data
```{r}
d.4stat <- 
  d%>%
  # remove setup data
  filter(Time>=0)%>%
  #make time an integer for correlation structure
  # by converting to days as time unit
  mutate(Time=as.integer(2*Time))%>%
  # for the lme model all fixed effects need to be factors
  mutate(time.fct=as.factor(Time))%>%
  mutate(seed.bank=as.factor(seed.bank))%>%
  mutate(phage=as.factor(phage))%>%
  mutate(cell.ml=log10(cell.ml))

d.4stat %>%
  ggplot(aes(x=Time, y=cell.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
        facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log10)")
```



# test data for homogeneity of variances
```{r}
# Levene's test with multiple independent variables
car::leveneTest(cell.ml ~ phage*seed.bank*line, data = d.4stat)
```


# RM-Anova with  corAR1 covariance structure

```{r}
rm.ar1 <- 
  d.4stat%>%
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * time.fct ,
    # lme(log10(cell.ml) ~ phage * seed.bank * time.fct ,
        random = ~1|flask, 
        correlation = corAR1(form = ~ Time | flask),
        data = .)

anova.lme(rm.ar1)
```


Looking at residuals  
```{r}
broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(Time,.resid)) +
  geom_hline(yintercept = 0)+
  geom_line(aes(color=line))+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(seed.bank,.resid)) +
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(fill=line),alpha=.5, position = position_dodge(width = .9))+
  geom_point(aes(color=line), position = position_dodge(width = .9), shape=21)+
  facet_wrap(~phage)+
  theme_cowplot()+panel_border()

broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(x = .resid, color = line))+
  geom_density()+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

```

```{r}
d.resid <- 
  broom.mixed::augment(rm.ar1)%>%
  select(.resid,phage,seed.bank,line)%>%
  group_by(phage,seed.bank,line)%>%
summarise(p.shapiro = shapiro.test(unlist(.resid))$p.value)%>%
  mutate(normal=if_else(p.shapiro>=0.05,TRUE,FALSE))%>%
  arrange(desc(p.shapiro))%>%
  print()

n.norm=sum(d.resid$normal)

broom.mixed::augment(rm.ar1)%>%
  select(.resid,phage,seed.bank,line)%>%
  group_by(phage,seed.bank,line)%>%
summarise(p.shapiro = shapiro.test(unlist(.resid))$p.value)%>%
  mutate(normal=if_else(p.shapiro>=0.05,TRUE,FALSE))%>%
  arrange(desc(p.shapiro))%>%
  ungroup()%>%
  mutate(rank=row_number())%>%
  ggplot(aes(rank, p.shapiro))+
  geom_col(aes(fill=seed.bank, color=phage),size=1)+
  geom_hline(yintercept = 0.05, color="red")+
  geom_vline(xintercept = n.norm+.5, color="blue")+
  scale_fill_grey()+
  scale_y_log10()+
  theme_cowplot()

```

More of the groups now have normally distributed residulas, but not all. 

# square root transformed data
```{r}
d.4stat <- 
  d%>%
  # remove setup data
  filter(Time>=0)%>%
  #make time an integer for correlation structure
  # by converting to days as time unit
  mutate(Time=as.integer(2*Time))%>%
  # for the lme model all fixed effects need to be factors
  mutate(time.fct=as.factor(Time))%>%
  mutate(seed.bank=as.factor(seed.bank))%>%
  mutate(phage=as.factor(phage))%>%
  mutate(cell.ml=sqrt(cell.ml))


d.4stat %>%
  ggplot(aes(x=Time, y=cell.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
        facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (sqrt)")+scale_y_log10()
```



# test data for homogeneity of variances
```{r}
# Levene's test with multiple independent variables
car::leveneTest(cell.ml ~ phage*seed.bank*line, data = d.4stat)
```


# RM-Anova with  corAR1 covariance structure

```{r}
rm.ar1 <- 
  d.4stat%>%
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * time.fct ,
    # lme(log10(cell.ml) ~ phage * seed.bank * time.fct ,
        random = ~1|flask, 
        correlation = corAR1(form = ~ Time | flask),
        data = .)

anova.lme(rm.ar1)
```


Looking at residuals  
```{r}
broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(Time,.resid)) +
  geom_hline(yintercept = 0)+
  geom_line(aes(color=line))+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(seed.bank,.resid)) +
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(fill=line),alpha=.5, position = position_dodge(width = .9))+
  geom_point(aes(color=line), position = position_dodge(width = .9), shape=21)+
  facet_wrap(~phage)+
  theme_cowplot()+panel_border()

broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(x = .resid, color = line))+
  geom_density()+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

```

```{r}
d.resid <- 
  broom.mixed::augment(rm.ar1)%>%
  select(.resid,phage,seed.bank,line)%>%
  group_by(phage,seed.bank,line)%>%
summarise(p.shapiro = shapiro.test(unlist(.resid))$p.value)%>%
  mutate(normal=if_else(p.shapiro>=0.05,TRUE,FALSE))%>%
  arrange(desc(p.shapiro))%>%
  print()

n.norm=sum(d.resid$normal)

broom.mixed::augment(rm.ar1)%>%
  select(.resid,phage,seed.bank,line)%>%
  group_by(phage,seed.bank,line)%>%
summarise(p.shapiro = shapiro.test(unlist(.resid))$p.value)%>%
  mutate(normal=if_else(p.shapiro>=0.05,TRUE,FALSE))%>%
  arrange(desc(p.shapiro))%>%
  ungroup()%>%
  mutate(rank=row_number())%>%
  ggplot(aes(rank, p.shapiro))+
  geom_col(aes(fill=seed.bank, color=phage),size=1)+
  geom_hline(yintercept = 0.05, color="red")+
  geom_vline(xintercept = n.norm+.5, color="blue")+
  scale_fill_grey()+
  scale_y_log10()+
  theme_cowplot()

```

#########################################################
# removing bi daily fluctuation

```{r}
# removing bi daily fluctuation
d.4stat <- 
  d%>%
  # remove setup data
  filter(Time>=0)%>%
  #make time an integer for correlation structure
  # by converting to days as time unit
  mutate(Time=as.integer(2*Time))%>%
  # for the lme model all fixed effects need to be factors
  mutate(time.fct=as.factor(Time))%>%
  mutate(seed.bank=as.factor(seed.bank))%>%
  mutate(phage=as.factor(phage))%>%
  mutate(cell.ml=log10(cell.ml))%>%
  # mutate(transfer=if_else(as.logical(Time%%2), "mid", "end"))%>%
  # group_by(transfer,phage,seed.bank)%>%
  # mutate(sdev =sd(cell.ml))%>%
  # ungroup()%>%
  # mutate(cell.ml=cell.ml-sdev)
  group_by(flask)%>%
  mutate(cell.ml=c(NA, diff(cell.ml)))%>%
  filter(!is.na(cell.ml))

# plot total cells of diff data
d.4stat%>%
  ggplot(aes(x=Time, y=cell.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
        facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (diff)")
```


# test data for homogeneity of variances
```{r}
# Levene's test with multiple independent variables
car::leveneTest(cell.ml ~ phage*seed.bank*line, data = d.4stat)
```


# RM-Anova with  corAR1 covariance structure

```{r}
rm.ar1 <- 
  d.4stat%>%
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * time.fct ,
    # lme(log10(cell.ml) ~ phage * seed.bank * time.fct ,
        random = ~1|flask, 
        correlation = corAR1(form = ~ Time | flask),
        data = .)

anova.lme(rm.ar1)
```


Looking at residuals  
```{r}
broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(Time,.resid)) +
  geom_hline(yintercept = 0)+
  geom_line(aes(color=line))+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(seed.bank,.resid)) +
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(fill=line),alpha=.5, position = position_dodge(width = .9))+
  geom_point(aes(color=line), position = position_dodge(width = .9), shape=21)+
  facet_wrap(~phage)+
  theme_cowplot()+panel_border()

broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(x = .resid, color = line))+
  geom_density()+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

```

```{r}
d.resid <- 
  broom.mixed::augment(rm.ar1)%>%
  select(.resid,phage,seed.bank,line)%>%
  group_by(phage,seed.bank,line)%>%
summarise(p.shapiro = shapiro.test(unlist(.resid))$p.value)%>%
  mutate(normal=if_else(p.shapiro>=0.05,TRUE,FALSE))%>%
  arrange(desc(p.shapiro))%>%
  print()

n.norm=sum(d.resid$normal)

broom.mixed::augment(rm.ar1)%>%
  select(.resid,phage,seed.bank,line)%>%
  group_by(phage,seed.bank,line)%>%
summarise(p.shapiro = shapiro.test(unlist(.resid))$p.value)%>%
  mutate(normal=if_else(p.shapiro>=0.05,TRUE,FALSE))%>%
  arrange(desc(p.shapiro))%>%
  ungroup()%>%
  mutate(rank=row_number())%>%
  ggplot(aes(rank, p.shapiro))+
  geom_col(aes(fill=seed.bank, color=phage),size=1)+
  geom_hline(yintercept = 0.05, color="red")+
  geom_vline(xintercept = n.norm+.5, color="blue")+
  scale_fill_grey()+
  scale_y_log10()+
  theme_cowplot()

```

More of the groups now have normally distributed residulas, but not all. 

#########################################################
# removing  fluctuation by decomposition

```{r, include=FALSE}
# initailise tibble for collecting decompsed data
d.decomp <- 
  tibble(Time=NA,
         line=NA,
         culture=NA,
         pop=NA,
         observed= NA,
         trend= NA,
         seasonal= NA,
         random= NA)


for (ln in levels(as.factor(d$line))){
  for (cult in levels(as.factor(d$culture))){
    for (pop in c("phage.ml","spore.ml","veg.ml","cell.ml")){
      
  #not all the lines have measurments of all the populations:
        #Ct lines have no phage
        #SN lines have no spores
  # the loop skips those combinations
      if (grepl("SN",cult) & pop=="spore.ml")
        next
      
      if (grepl("Ct",cult) & pop=="phage.ml")
        next
      
  tmp <- 
    d%>%
  filter(Time>=0)%>%
  filter(culture==cult)%>%
  filter(line==ln)%>%
  select(all_of(pop))%>%
ts(.[,pop], start=0, end=14,frequency = 2 )%>%
  decompose()

  # get trend data
  d.decomp <- 
  tibble(Time=seq(0,14,0.5),
         line=ln,
         culture=cult,
         pop=pop,
         observed= as.vector(tmp$x),
         trend= as.vector(tmp$trend),
         seasonal= as.vector(tmp$seasonal),
         random= as.vector(tmp$random))%>%
    bind_rows(d.decomp,.)
  tmp$trend%>%as.vector()
  
  #plot
    #   ggplot2::autoplot(tmp)+
    #     theme_bw()+
    # ggtitle(paste(cult,ln,pop))+
    #     ggsave(filename = here("plots","deomposition",paste(pop,cult,ln,".pdf", sep = "_")),width = 8, height = 6)
  
    }

  }

}

#remove first row used to initialise
d.decomp <- d.decomp[-1,]

#delete temporary variables
rm(tmp,cult,ln,pop)

#return to wide format
d.decomp.w <- 
  d.decomp%>%
  filter(!is.na(trend))%>%
  select(-observed,-seasonal,-random)%>%
  pivot_wider( names_from="pop", values_from =trend)

# #add columns on host strain,phage treatment and seed bank as encoded in culture name
d.decomp.w <-
   d.decomp.w%>%
   mutate(host=case_when(grepl("W", culture) ~ "WT",
                         grepl("SN", culture) ~ "dSpoIIE",
                         grepl("dS", culture) ~ "dSpoIIE"))%>%

  mutate(phage=case_when(grepl("O", culture) ~ "SPO1",
                         grepl("Ct", culture) ~ "no_Phage",
                         grepl("ANC", line) ~ "no_Phage"))%>%

  mutate(seed.bank=case_when(grepl("SN", culture) ~ "none",
                              grepl("WS", culture) ~ "short",
                              grepl("WL", culture) ~ "long"))%>%
   mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))%>%
  #add flask to be used as grouping variable
  mutate(flask=paste(culture,line, sep = "-"))%>%
  #remove setup data
  filter(Time>=0)


```

```{r}
# removing bi daily fluctuation
d.4stat <- 
  d.decomp.w %>%
  # remove setup data
  filter(Time>=0)%>%
  #make time an integer for correlation structure
  # by converting to days as time unit
  mutate(Time=as.integer(2*Time))%>%
  # for the lme model all fixed effects need to be factors
  mutate(time.fct=as.factor(Time))%>%
  mutate(seed.bank=as.factor(seed.bank))%>%
  mutate(phage=as.factor(phage))%>%
  mutate(cell.ml=log10(cell.ml))

# plot total cells of diff data
d.4stat%>%
  ggplot(aes(x=Time, y=cell.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
        facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (decomp)")
```


# test data for homogeneity of variances
```{r}
# Levene's test with multiple independent variables
car::leveneTest(cell.ml ~ phage*seed.bank*line, data = d.4stat)
```


# RM-Anova with  corAR1 covariance structure

```{r}
rm.ar1 <- 
  d.4stat%>%
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * time.fct ,
    # lme(log10(cell.ml) ~ phage * seed.bank * time.fct ,
        random = ~1|flask, 
        correlation = corAR1(form = ~ Time | flask),
        data = .)

anova.lme(rm.ar1)
```


Looking at residuals  
```{r}
broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(Time,.resid)) +
  geom_hline(yintercept = 0)+
  geom_line(aes(color=line))+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(seed.bank,.resid)) +
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(fill=line),alpha=.5, position = position_dodge(width = .9))+
  geom_point(aes(color=line), position = position_dodge(width = .9), shape=21)+
  facet_wrap(~phage)+
  theme_cowplot()+panel_border()

broom.mixed::augment(rm.ar1)%>%
  ggplot(aes(x = .resid, color = line))+
  geom_density()+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

```

```{r}
d.resid <- 
  broom.mixed::augment(rm.ar1)%>%
  select(.resid,phage,seed.bank,line)%>%
  group_by(phage,seed.bank,line)%>%
summarise(p.shapiro = shapiro.test(unlist(.resid))$p.value)%>%
  mutate(normal=if_else(p.shapiro>=0.05,TRUE,FALSE))%>%
  arrange(desc(p.shapiro))%>%
  print()

n.norm=sum(d.resid$normal)

broom.mixed::augment(rm.ar1)%>%
  select(.resid,phage,seed.bank,line)%>%
  group_by(phage,seed.bank,line)%>%
summarise(p.shapiro = shapiro.test(unlist(.resid))$p.value)%>%
  mutate(normal=if_else(p.shapiro>=0.05,TRUE,FALSE))%>%
  arrange(desc(p.shapiro))%>%
  ungroup()%>%
  mutate(rank=row_number())%>%
  ggplot(aes(rank, p.shapiro))+
  geom_col(aes(fill=seed.bank, color=phage),size=1)+
  geom_hline(yintercept = 0.05, color="red")+
  geom_vline(xintercept = n.norm+.5, color="blue")+
  scale_fill_grey()+
  scale_y_log10()+
  theme_cowplot()

```

More of the groups now have normally distributed residulas, but not all. 