---
title: "RM-ANOVA for coevolution with a seed bank"
subtitle: "A. Organize the data"
output:
  pdf_document: default
  html_notebook:
    code_folding: hide
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(renv)
# #install to project library the required packages
# renv::restore()
# # Initialize lock file when first setting repo
# renv::init()

library(here)
# # set root at project file ("coevolution-ts"). 
# # Done once at initial setup.
# here::set_here()

library(tidyverse)
library(cowplot)

# save the state of the project library to the lockfile (called renv.lock),
# renv::snapshot()

```

## Population density data  
I have previously consolidated all the data on population density collected for this experiment. This includes:

| Population | What it is | Method of measurement |
| - | - | - |
|spore / mL | host spore density | flow-cytometry: low SYBR population |
|veg / mL | host vegetative cell density | flow-cytometry: high SYBR population |
| cell /mL | total host cell density (spore+veg) | flow-cytometry
| phage/ mL | total SPO1 phage density (free phage) | qPCR (SYBR)  with phage specific primers and lysate serial dilution as standard | 

```{r load data, echo=FALSE, message=FALSE}
d <- read_csv(here("data","coevolution_pop_density.csv") )

#add columns on host strain and phage treatment  as encoded in culture name
d <- 
  d%>%
  # host 
  mutate(host=case_when(grepl("W", culture) ~ "WT",
                         grepl("SN", culture) ~ "dSpoIIE",
                         grepl("dS", culture) ~ "dSpoIIE"))%>%
  # phage
  mutate(phage=case_when(grepl("O", culture) ~ "SPO1",
                         grepl("Ct", culture) ~ "no_Phage",
                         grepl("ANC", line) ~ "no_Phage"))%>%
  #seed bank
  mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))%>%
  #add flask to be used as grouping variable
  mutate(flask=paste(culture,line, sep = "-"))%>%
  #remove setup data (prior to phage infection)
  filter(Time>=0)

```

```{r}
# plot raw data
d%>%
    pivot_longer(cols=contains(".ml"), names_to="pop",values_to="num.ml")%>%
    ggplot(aes(x=Time, y=num.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      facet_grid(pop~phage)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 1e10))
```

In the first few days of the experiment cell numbers dropped below detection limit in 2 of 3 SNO lines. Missing data cannot be analyzed in RM-ANOVA. I will assign these points the detection limit values.

Points with missing data: 
```{r}
d%>%
  filter(cell.ml<10) 
```
 
 These are all veg cells:

```{r}
d%>%
  filter(veg.ml<10)
```
---  

**Flow-cytometry detection limit**

The flow cytometer stop conditions were either $50\mu L$ or 50,000 events, whatever came first.  
In addition, to account for false asignments by noise I required that a gated population have at least 100 events to be taken into account. So the minimal detected population would have 100 cells in $50\mu L$ of analyzed liquid. At this poimt of the experiment we were analyzing x100 diluted samples.

```{r}
detection.limit <- 100 * (1000/50) * 100
```

100 events X $1000/50\mu L$ X100 = `r detection.limit` 

I will use $1X10^5$, half the detection limit dor missing value points
 

```{r}
d <- d%>%
  mutate(cell.ml=if_else(cell.ml>10, cell.ml,detection.limit/2))%>%
  mutate(veg.ml=if_else(veg.ml>10, veg.ml,detection.limit/2))


# plot total cells of raw data
d%>%
    ggplot(aes(x=Time, y=cell.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
  geom_hline(yintercept = detection.limit, color="red")+
  geom_label(label="detection limit", x=10, y=log10(detection.limit), color="red")+
      facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 2e9))
```



```{r}
# Save the data to be used for statistical analyes
write_csv(d,here("data","organized_pop_density.csv"))

```

