---
title: "RM-ANOVA for coevolution with a seed bank"
subtitle: "Total cell density"
output:
  pdf_document: default
  html_notebook:
    code_folding: hide
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(renv)
# #install to project library the required packages
# renv::restore()
# # Initialize lock file when first setting repo
# renv::init()

library(here)
# # set root at project file ("coevolution-ts"). 
# # Done once at initial setup.
# here::set_here()

library(tidyverse)
library(cowplot)


library(nlme)
library(car)
# library(ggpubr)
library(broom)
library(broom.mixed)
library(emmeans)

# save the state of the project library to the lockfile (called renv.lock),
# renv::snapshot()

```

Load organized data  
```{r, echo=FALSE, message=FALSE}
d <- read_csv(here("data","organized_pop_density.csv"), )

```

### Adjusting variable types for lme

In the lme models used below the fixed effects need to be specified as factors. In this experiment these are: *Phage, seed.bank and time*. However for the auto correlation specification time needs to be specified as an integer. To fulfill both requirements we use  the experimental day rather than transfer as the time unit, since samples taken once a day but twice per transfer. This will be simply *time x 2*. From that we make a separate variable which will be the factor of the time.

```{r}
d <- 
  d%>%
  #make time an integer for correlation structure
  # by converting to days as time unit
  mutate(Time=as.integer(2*Time))%>%
  # for the lme model all fixed effects need to be factors
  mutate(time.fct=as.factor(Time))%>%
  mutate(phage=as.factor(phage))%>%
  mutate(seed.bank=as.factor(seed.bank))%>%
  #ajust factor order for seed bank
  mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))
```

Select response variable to be analyzed  

```{r}
var.response <- "cell/ml" 

d <- d %>%
  mutate(response=cell.ml )
```

Here we analyze **`r var.response` **.

```{r}
d%>%
ggplot(aes(x=Time, y=response))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
        facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab(var.response)
```


### Test data for homogeneity of variances

This is an assumption of ANOVA tests.

Based on : https://www.datanovia.com/en/lessons/homogeneity-of-variance-test-in-r/
Using "Levene’s test" that according to website is the most commonly used test for this purpose. This test has a null hypothesis of equal variance. So getting P>0.05 suggests homogenic variance.  

```{r}
# Levene's test with multiple independent variables
car::leveneTest(response ~ phage*seed.bank*line, data = d)
```

**The data does not fulfill the assumption of equal variance across test groups.**


## Log transform the data  


```{r}
d <- d%>%
  mutate(log.response=log10(response))


d%>%
ggplot(aes(x=Time, y=log.response))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
        facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab(paste(var.response,"(log10)"))
```


### Test transformed data for homogeneity of variances

```{r}
# Levene's test with multiple independent variables
car::leveneTest(log.response ~ phage*seed.bank*line, data = d)
```

**The data still does not fulfill the assumption of equal variance across test groups.**


# Box-Cox transformation

>`powerTransform` uses the maximum likelihood-like approach of Box and Cox (1964) to select a transformation of a univariate or multivariate response for normality, linearity and/or constant variance.  
>(help page for `car::powerTransform`)


```{r}
# Multivariate transformation to normality within levels of 'htype'
bx.cx <- powerTransform(response ~ phage*seed.bank*line, d)
summary(bx.cx)
```

*Tranformation is required, but not a simple log transformation*

Plot the profile log-likelihood for Box-Cox transformations.  
```{r}
  with(d, boxCox(cell.ml ~ phage*seed.bank*line, 
                 lambda = seq(-1, bx.cx$roundlam+3, by = 0.1),
                 family="bcPower"))
```

Transform using Box-Cox $\lambda$ (rounded). 
```{r}
d <- d%>%
  mutate(bxcx.response=bcPower(cell.ml, bx.cx$roundlam))

d%>%
ggplot(aes(x=Time, y=bxcx.response))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
        facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab(paste(var.response,"(Box-Cox transformed)"))

```

### Test transformed data for homogeneity of variances

```{r}
# Levene's test with multiple independent variables
car::leveneTest(bxcx.response ~ phage*seed.bank*line, data = d)
```

**The data now fulfills the assumption of equal variance across test groups.**

# RM-ANOVA model selection

To account for time series auto-correlation we will specify correlation structure to the lme model. We will test various corARMA correlation structures and choose by lowest AIC. This is a combined  auto-regressive model, AR(p), and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC). Since we have a seasonality of lag 2 due to transfer we will look at lag up to 2 for both AR and MA. Note that a corARMA(p=1, q=0) is the same as corAR1.  

```{r arma, message = FALSE, warning = FALSE}
# initalise empty list to save models
l.rm <- list()

# initalise empty table to collect model data
pq.aic <- tibble()


  
for(Q in c(0:2)){ 
  for (P in c(0:2)){ 
    
    #skip corARMA(0,0)
    if (P==0 & Q==0) next
    
    #run model
    cur.model <- d%>%
      lme(bxcx.response ~ phage * seed.bank * time.fct , 
          random = ~1|flask, 
          correlation = corARMA(form = ~ Time | flask, p=P,q=Q),
            data = .)
    l.rm[[paste0("ARMA_P",P,"Q",Q)]] <- cur.model
    
    cur.tbl <- broom.mixed::glance(cur.model)%>%
      mutate(p=P)%>%
      mutate(q=Q)

      
      pq.aic <- bind_rows(pq.aic,cur.tbl)
    
  }
}

# get list order by AIC
pq.aic <- 
  pq.aic%>%arrange(AIC)%>%
  # model name
  mutate(arma=paste0("ARMA_P",p,"Q",q))

# save the best model
m.best <- l.rm[[pq.aic$arma[1]]]
best <- paste0("corARMA(",pq.aic$p[1],",",pq.aic$q[1],")")

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,AIC))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)",paste("best model uses", best))

# compare best 5 models by AIC order
t.models <- anova(l.rm[[pq.aic$arma[5]]],
                  l.rm[[pq.aic$arma[4]]],
                  l.rm[[pq.aic$arma[3]]],
                  l.rm[[pq.aic$arma[2]]],
                  l.rm[[pq.aic$arma[1]]])

t.models%>%
  tibble()%>%
  select(-call)%>%
  mutate(arma=pq.aic$arma[1:5])

```


# best model is `r best`

Though it is not significantly better than the next 2 models.  

**Results of selected model**

```{r}
# best model data
pq.aic%>%
  slice_min(AIC)
#display best model results
  anova(m.best)

  
```

As expected time is very significant at all levels. This includes also a significant 3-way interaction phage x seed-bank x time is detected.  
Phage too is strongly significant as a main effect and in interaction with time. Phage X seed-bank is on border of significance.  
Seed-bank is significant as main effect

There is evidence to suggest that the seed bank treatment is influencing population dynamics in a phage dependent manner. Before looking into the potential drivers of this effect we evaluate the model.


### How sensitive would the result be to model selected?

looking at different correlation structures.

```{r}

# all model results to tibble
all.models <- 
  map(l.rm, anova.lme) %>%
  map(rownames_to_column)%>%
  bind_rows(.id = "arma")%>%
  rename(effect=rowname)


all.models%>%
  mutate(effect=fct_inorder(effect))%>%
  ggplot(aes(x=effect, y=`p-value`))+
  geom_hline(yintercept = c(0.05,0.01,1e-3), color="pink")+
  geom_jitter(aes(fill=effect),width = 0.2, height = 0,
              shape=21, size=2, show.legend = F)+
  coord_flip()+
  theme_cowplot()

```

The results are pretty similar across the models. However for the **phageXseed.bank** interaction the effect is borderline significant with 4/6 of the models having a P<0.05 for that interaction. 


#### Distribution of model residuals

For the best model.

```{r,fig.height=10, fig.width=15}

p1 <- 
  #qqplot by seed bank
  qqnorm(m.best,~ resid(., type = "p")|seed.bank, abline = c(0, 1))

p2 <-
  # standardized residuals versus fitted values by seed.bank
plot(m.best, resid(., type = "p") ~ fitted(.) | seed.bank, abline = 0)

p3 <-
broom.mixed::augment(m.best)%>%
  ggplot(aes(Time,.resid)) +
  geom_hline(yintercept = 0)+
  geom_line(aes(color=line))+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

p4 <-
broom.mixed::augment(m.best)%>%
  ggplot(aes(seed.bank,.resid)) +
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(fill=line),alpha=.5, position = position_dodge(width = .9))+
  geom_point(aes(color=line), position = position_dodge(width = .9), shape=21)+
  facet_wrap(~phage)+
  theme_cowplot()+panel_border()



plot_grid(p1,p2,p3,p4, nrow = 2, labels = 'auto')

```

conclusions:  
a. model residuals are very close to normal distribution. Sign of good fit.  
b. The residual are evenly distributed around 0, suggesting equal variance.  
c. Equal residual variance holds across time.  
d. Equal residual variance holds across experimental units (flasks).  

*A single clear deviation is the no-seed-bank L2 flask.* 

# post hoc  
What in the phage X seed-bank interaction is causing a significant effect on population dynamics?  
We analyze the *estimated marginal means (EMM)*, following examples from: https://cran.r-project.org/web/packages/emmeans/vignettes/interactions.html

### Visualize EMM of interactions
```{r}
# plot 
emmip(m.best, phage~seed.bank)+
  theme_cowplot()
```

Plot shows that the no-seed-bank in the presence of phage is likely the treatment interaction standing out. We compare seed banks treatments separately for phage treatments.

```{r}
emm_sb.ph <- emmeans(m.best, pairwise ~ seed.bank | phage)

emm_sb.ph$contrasts

plot(emm_sb.ph)+
  theme_cowplot()+
  panel_border()


```

Indeed in the phage infected treatments we see that the populations differ between no-seed-bank lines and each of the with seed-bank treatments. But there is no difference between long and short seed banks. The trend is that the absence of a seed.bank results in lower cell densities.  
We also saw an interaction with time. We next compare the seed-bank treatments across time, separating the phage treatments, and focusing on the phage treated samples.

```{r}

coevo.emm <- emmeans(m.best, ~ seed.bank * phage * time.fct)
coevo.pairs <- pairs(coevo.emm, simple="seed.bank")%>%
  tidy

p1 <- 
  coevo.pairs%>%
  # focus on phage infected
  filter(phage=="SPO1")%>%
  # make time continuous for plotting.
  mutate(day=as.numeric(time.fct))%>%
  # arrange panel order
  mutate(contrast=fct_relevel(contrast, "long - short", after = 0))%>%
  #plot
  ggplot(aes(x=day, y=estimate))+
    #add 0 line
  geom_hline(yintercept = 0, color="red")+
  geom_point()+
  geom_line()+
  facet_grid(contrast~.)+
  theme_cowplot()+
  panel_border()

p2 <- coevo.pairs%>%
  # focus on phage infected
  filter(phage=="SPO1")%>%
  # make time continuous for plotting.
  mutate(day=as.numeric(time.fct))%>%
  # arrange panel order
  mutate(contrast=fct_relevel(contrast, "long - short", after = 0))%>%
  #plot
  ggplot(aes(x=day, y=adj.p.value))+
    #add 0.05 signifcance thrshold
  geom_hline(yintercept = 0.05, color="red")+
  geom_point()+
  geom_line()+
  facet_grid(contrast~.)+
  theme_cowplot()+
  panel_border()



plot_grid(p1,p2, labels = "auto") 
```

The "absence of seed bank" effect on population density is seen in days 4-12 of the experiment. In that period a significant reduction in cell density is seen. However after that time the effect of no-seed-bank is lost, until the last samples were as a trend the population density is becoming lower in absence of seed bank and that hits the significance threshold on the last sample day. 

# Summary

The seed bank treatment alters population dynamics (cell density over time) in phage infected cultures only. Populations coevolving with phages in the absence of seed bank show a significant reduction in cell densities, most notably in the short to medium time frame. This is to say that the presence of a seed bank acts as a buffer that moderates the reduction in host population density caused by phage infection. The results suggest however that other mechanisms (presumably host evolution) can eventually enable the host to achieve the same resilience against phage-induced decimation of its population , but these may be less stable in the long run. 