---
title: "RM-ANOVA for coevolution with a seed bank"
subtitle: "Phage density"
output:
  pdf_document: default
  html_notebook:
    code_folding: hide
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(renv)
# #install to project library the required packages
# renv::restore()
# # Initialize lock file when first setting repo
# renv::init()

library(here)
# # set root at project file ("coevolution-ts"). 
# # Done once at initial setup.
# here::set_here()

library(tidyverse)
library(cowplot)


library(nlme)
library(car)
library(ggpubr)
library(broom)
library(broom.mixed)

# save the state of the project library to the lockfile (called renv.lock),
# renv::snapshot()

```

Load organized data  
```{r, echo=FALSE, message=FALSE}
d <- read_csv(here("data","organized_pop_density.csv"), )

```

### Adjusting variable types for lme

In the lme models used below the fixed effects need to be specified as factors. In this experiment these are: *Phage, seed.bank and time*. However for the auto correlation specification time needs to be specified as an integer. To fulfill both requirements we use  the experimental day rather than transfer as the time unit, since samples taken once a day but twice per transfer. This will be simply *time x 2*. From that we make a separate variable which will be the factor of the time.

```{r}
d <- 
  d%>%
  #make time an integer for correlation structure
  # by converting to days as time unit
  mutate(Time=as.integer(2*Time))%>%
  # for the lme model all fixed effects need to be factors
  mutate(time.fct=as.factor(Time))%>%
  mutate(phage=as.factor(phage))%>%
  mutate(seed.bank=as.factor(seed.bank))%>%
  #ajust factor order for seed bank
  mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))
```

Select reponse variable to be analyzed  

```{r}
var.response <- "phage/ml" 

d <- d %>%
  mutate(response=phage.ml )%>%
  # remove no phage data
  filter(phage=="SPO1")%>%
  # remove time of infection (T=0) the chnage in that first peios is an oulier to the data
  filter(Time>0)
```

Here we analyze **`r var.response` **.

```{r}
d%>%
ggplot(aes(x=Time, y=response))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
        # facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab(var.response)
```


### Test data for homogeneity of variances

This is an assumption of ANOVA tests.

Based on : https://www.datanovia.com/en/lessons/homogeneity-of-variance-test-in-r/
Using "Levene’s test" that according to website is the most commonly used test for this purpose. This test has a null hypothesis of equal variance. So getting P>0.05 suggests homogenic variance.  

```{r}
# Levene's test with multiple independent variables
car::leveneTest(response ~ phage*seed.bank*line, data = d)
```

The data fulfills the assumption of equal variance across test groups, with no need for transformation. I will log transform just for plotting:

**Log transform the data**

```{r}
d <- d%>%
  mutate(log.response=log10(response))


d%>%
  ggplot(aes(x=Time, y=log.response))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
        # facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab(paste(var.response,"(log10)"))
```


### Test transformed data for homogeneity of variances

```{r}
# Levene's test with multiple independent variables
car::leveneTest(log.response ~ phage*seed.bank*line, data = d)
```


# Box-Cox transformation

`powerTransform` uses the maximum likelihood-like approach of Box and Cox (1964) to select a transformation of a univariate or multivariate response for normality, linearity and/or constant variance. (help page for `car::powerTransform`)

we will run the Cox-Cox transformation procedure here too, since this also tests the hypothesis that no transformation is needed ($\lambda =1$).

```{r}
# Multivariate transformation to normality within levels of 'htype'
bx.cx <- powerTransform(response ~ seed.bank*line, d)
summary(bx.cx)
```

The null hypothesis of no transformation needed ($\lambda =1$) is rejected, and so we will transform the data using $\lambda$ suggested by Box-Cox.

Plot the profile log-likelihood for Box-Cox transformations .  
```{r}
  with(d, boxCox(response ~ seed.bank*line, 
                 lambda = seq(-1, bx.cx$roundlam+3, by = 0.1),
                 family="bcPower"))
```

Transform using Box-Cox $\lambda$ (rounded). 
```{r}
d <- d%>%
  mutate(bxcx.response=bcPower(response, bx.cx$roundlam))

d%>%
ggplot(aes(x=Time, y=bxcx.response))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
        facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab(paste(var.response,"(Box-Cox transformed)"))

```

### Test transformed data for homogeneity of variances

```{r}
# Levene's test with multiple independent variables
car::leveneTest(bxcx.response ~ phage*seed.bank*line, data = d)
```

The data still fulfills the assumption of equal variance across test groups.

# RM-ANOVA model selection

To account for time series auto-correlation we will specify correlation structure to the lme model. We will test various corARMA correlation structures and choose by lowest AIC. This is a combined  auto-regressive model, AR(p), and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC). Since we have a seasonality of lag 2 due to transfer we will look at lag up to 2 for both AR and MA. Note that a corARMA(p=1, q=0) is the same as corAR1.  

```{r arma, message = FALSE, warning = FALSE}
# initalise empty list to save models
l.rm <- list()

# initalise empty table to collect model data
pq.aic <- tibble()


  
for(Q in c(0:2)){ 
  for (P in c(0:2)){ 
    
    #skip corARMA(0,0)
    if (P==0 & Q==0) next
    
    #run model
    cur.model <- d%>%
      lme(bxcx.response ~  seed.bank * time.fct , 
          random = ~1|flask, 
          correlation = corARMA(form = ~ Time | flask, p=P,q=Q),
            data = .)
    l.rm[[paste0("ARMA_P",P,"Q",Q)]] <- cur.model
    
    cur.tbl <- broom.mixed::glance(cur.model)%>%
      mutate(p=P)%>%
      mutate(q=Q)

      
      pq.aic <- bind_rows(pq.aic,cur.tbl)
    
  }
}

# get list order by AIC
pq.aic <- 
  pq.aic%>%arrange(AIC)%>%
  # model name
  mutate(arma=paste0("ARMA_P",p,"Q",q))

best <- paste0("corARMA(",pq.aic$p[1],",",pq.aic$q[1],")")

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,AIC))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)",paste("best model uses", best))

# compare best 5 models by AIC order
t.models <- anova(l.rm[[pq.aic$arma[5]]],
                  l.rm[[pq.aic$arma[4]]],
                  l.rm[[pq.aic$arma[3]]],
                  l.rm[[pq.aic$arma[2]]],
                  l.rm[[pq.aic$arma[1]]])

t.models%>%
  tibble()%>%
  select(-call)%>%
  mutate(arma=pq.aic$arma[1:5])

```


# best model is `r best`

Though it is not significantly better than any of the other models.  
This is the same correlation-strucure chosen for the total cells.

**Results of selected model**

```{r}
# best model data
pq.aic%>%
  slice_min(AIC)
#display best model results
  anova(l.rm[[pq.aic$arma[1]]])
  
```





### How sensitive would the result be to model selected?
```{r}

# putl all model results in tibble
all.models <- 
  map(l.rm, anova.lme) %>%
  map(rownames_to_column)%>%
  bind_rows(.id = "arma")%>%
  rename(effect=rowname)


all.models%>%
  mutate(effect=fct_inorder(effect))%>%
  ggplot(aes(x=effect, y=`p-value`))+
  geom_hline(yintercept = c(0.05,0.01,1e-3), color="pink")+
  geom_jitter(aes(fill=effect),width = 0.2, height = 0,
              shape=21, size=2, show.legend = F)+
  coord_flip()+
  theme_cowplot()

```

The results are simmilar across the models at least in all models leading to saame conclusions on significance.


#### Distribution of model residuals

For the best model.

```{r,fig.height=10, fig.width=15}


p1 <- ggqqplot(residuals(l.rm[[pq.aic$arma[1]]]))

p2 <-
broom.mixed::augment(l.rm[[pq.aic$arma[1]]])%>%
  ggplot(aes(Time,.resid)) +
  geom_hline(yintercept = 0)+
  geom_line(aes(color=line))+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

p3 <-
broom.mixed::augment(l.rm[[pq.aic$arma[1]]])%>%
  ggplot(aes(seed.bank,.resid)) +
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(fill=line),alpha=.5, position = position_dodge(width = .9))+
  geom_point(aes(color=line), position = position_dodge(width = .9), shape=21)+
  facet_wrap(~phage)+
  theme_cowplot()+panel_border()

p4 <-
broom.mixed::augment(l.rm[[pq.aic$arma[1]]])%>%
  ggplot(aes(x = .resid, color = line))+
  geom_density()+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

plot_grid(p1,p2,p3,p4, nrow = 2, labels = 'auto')

```


