---
title: "RM-ANOVA for coevolution with a seed bank"
subtitle: "Total cell density"
output:
  pdf_document: default
  html_notebook:
    code_folding: hide
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(renv)
# #install to project library the required packages
# renv::restore()
# # Initialize lock file when first setting repo
# renv::init()

library(here)
# # set root at project file ("coevolution-ts"). 
# # Done once at initial setup.
# here::set_here()

library(tidyverse)
library(cowplot)


library(nlme)
library(car)
library(emmeans)
library(broom)
library(broom.mixed)

# save the state of the project library to the lockfile (called renv.lock),
# renv::snapshot()

```

Load organized data  
```{r, echo=FALSE, message=FALSE}
d <- read_csv(here("data","organized_pop_density.csv"), )

```

### Adjusting variable types for lme

In the lme models used below the fixed effects need to be specified as factors. In this experiment these are: *Phage, seed.bank and time*. However for the auto correlation specification time needs to be specified as an integer. To fulfill both requirements we use  the experimental day rather than transfer as the time unit, since samples taken once a day but twice per transfer. This will be simply *time x 2*. From that we make a separate variable which will be the factor of the time.

```{r}
d <- 
  d%>%
  #make time an integer for correlation structure
  # by converting to days as time unit
  mutate(Time=as.integer(2*Time))%>%
  # for the lme model all fixed effects need to be factors
  mutate(time.fct=as.factor(Time))%>%
  mutate(phage=as.factor(phage))%>%
  mutate(seed.bank=as.factor(seed.bank))%>%
  #ajust factor order for seed bank
  mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))
```

Select response variable to be analyzed  

```{r}
var.response <- "%spore/ml" 

d <- d %>%
  #remove non-sporulating host sample
  filter(host=="WT")%>%
  mutate(response=100*spore.ml/cell.ml )
```

Here we analyze **`r var.response` **.

```{r}
d%>%
ggplot(aes(x=Time, y=response))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
        facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab(var.response)
```


### Test data for homogeneity of variances

This is an assumption of ANOVA tests.

Based on : https://www.datanovia.com/en/lessons/homogeneity-of-variance-test-in-r/
Using "Levene’s test" that according to website is the most commonly used test for this purpose. This test has a null hypothesis of equal variance. So getting P>0.05 suggests homogenic variance.  

```{r}
# Levene's test with multiple independent variables
car::leveneTest(response ~ phage*seed.bank*line, data = d)
```

The data does not fulfill the assumption of equal variance across test groups.


## Log transform the data

```{r}
d <- d%>%
  mutate(log.response=log10(response))


d%>%
ggplot(aes(x=Time, y=log.response))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
        facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab(paste(var.response,"(log10)"))
```


### Test transformed data for homogeneity of variances

```{r}
# Levene's test with multiple independent variables
car::leveneTest(log.response ~ phage*seed.bank*line, data = d)
```

The data  fulfills the assumption of equal variance across test groups, at a borderline p-value.


# Box-Cox transformation

`powerTransform` uses the maximum likelihood-like approach of Box and Cox (1964) to select a transformation of a univariate or multivariate response for normality, linearity and/or constant variance. (help page for `car::powerTransform`)

```{r}
# Multivariate transformation to normality within levels of 'htype'
bx.cx <- powerTransform(response ~ phage*seed.bank*line, d)
summary(bx.cx)
```
The null hypothesis of no transformation needed ($\lambda =1$) is **NOT** rejected and so we  proceed to **do the analysis on non transformed data**. 


Plot the profile log-likelihood for Box-Cox transformations .  
```{r}
  with(d, boxCox(response ~ phage*seed.bank*line, 
                 lambda = seq(-1, bx.cx$roundlam+3, by = 0.1),
                 family="bcPower"))
```



# RM-ANOVA model selection

To account for time series auto-correlation we will specify correlation structure to the lme model. We will test various corARMA correlation structures and choose by lowest AIC. This is a combined  auto-regressive model, AR(p), and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC). Since we have a seasonality of lag 2 due to transfer we will look at lag up to 2 for both AR and MA. Note that a corARMA(p=1, q=0) is the same as corAR1.  

```{r arma, message = FALSE, warning = FALSE}
# initalise empty list to save models
l.rm <- list()

# initalise empty table to collect model data
pq.aic <- tibble()


  
for(Q in c(0:2)){ 
  for (P in c(0:2)){ 
    
    #skip corARMA(0,0)
    if (P==0 & Q==0) next
    
    # corARMA(1,2) had singular convergence
    if (P==1 & Q==2) next
    
    #run model
    cur.model <- d%>%
      lme(response ~ phage * seed.bank * time.fct , 
          random = ~1|flask, 
          correlation = corARMA(form = ~ Time | flask, p=P,q=Q),
            data = .)
    l.rm[[paste0("ARMA_P",P,"Q",Q)]] <- cur.model
    
    cur.tbl <- broom.mixed::glance(cur.model)%>%
      mutate(p=P)%>%
      mutate(q=Q)

      
      pq.aic <- bind_rows(pq.aic,cur.tbl)
    
  }
}

# get list order by AIC
pq.aic <- 
  pq.aic%>%arrange(AIC)%>%
  # model name
  mutate(arma=paste0("ARMA_P",p,"Q",q))

# save the best model
m.best <- l.rm[[pq.aic$arma[1]]]
best <- paste0("corARMA(",pq.aic$p[1],",",pq.aic$q[1],")")

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,AIC))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)",paste("best model uses", best))

# compare best 5 models by AIC order
t.models <- anova(l.rm[[pq.aic$arma[5]]],
                  l.rm[[pq.aic$arma[4]]],
                  l.rm[[pq.aic$arma[3]]],
                  l.rm[[pq.aic$arma[2]]],
                  l.rm[[pq.aic$arma[1]]])

t.models%>%
  tibble()%>%
  select(-call)%>%
  mutate(arma=pq.aic$arma[1:5])

```


# best model is `r best`

This model is significantly better than the other models.  

**Results of selected model**

```{r}
# best model data
pq.aic%>%
  slice_min(AIC)
#display best model results
  anova(m.best)
  
```

The seed-bank treatment is found as a significant effect in interaction with time, and nearly significant in interaction with phage. phage and time both are significant effect as is their interaction. Before looking into the potential drivers of these effects we evaluate the model. 


### How sensitive would the result be to model selected?  
```{r}

# putl all model results in tibble
all.models <- 
  map(l.rm, anova.lme) %>%
  map(rownames_to_column)%>%
  bind_rows(.id = "arma")%>%
  rename(effect=rowname)

arma.best <- paste0("ARMA_P",pq.aic$p[1],"Q",pq.aic$q[1])
selected <- 
  all.models%>%
  filter(arma==arma.best)


all.models%>%
  mutate(effect=fct_inorder(effect))%>%
  ggplot(aes(x=effect, y=`p-value`))+
   geom_point(data = selected, color="black", size=10, shape="|")+
  geom_hline(yintercept = c(0.05,0.01,1e-3), color="pink")+
  geom_jitter(aes(fill=effect),width = 0.2, height = 0,
              shape=21, size=2, show.legend = F)+
 
  coord_flip()+
  theme_cowplot()
```

The line marks the p-values of the selected model, and the points show p-values of all models tested (including best). There is variance in p-values associated with seed bank treatment as main effect, in interaction with phage and in 3-way interaction. However the results lead to similar conclusions across the models, with the exception of phage X seed-bank interaction which hovers around the .05 significance threshold with some models yielding lower than threshold p-values.


#### Distribution of model residuals

For the best model.

```{r,fig.height=10, fig.width=15}
p1 <- 
  #qqplot by seed bank
  qqnorm(m.best,~ resid(., type = "p")|seed.bank, abline = c(0, 1))

p2 <-
  # standardized residuals versus fitted values by seed.bank
plot(m.best, resid(., type = "p") ~ fitted(.) | seed.bank, abline = 0)

p3 <-
broom.mixed::augment(m.best)%>%
  ggplot(aes(Time,.resid)) +
  geom_hline(yintercept = 0)+
  geom_line(aes(color=line))+
  facet_grid(seed.bank~phage)+
  theme_cowplot()+panel_border()

p4 <-
broom.mixed::augment(m.best)%>%
  ggplot(aes(seed.bank,.resid)) +
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(fill=line),alpha=.5, position = position_dodge(width = .9))+
  geom_point(aes(color=line), position = position_dodge(width = .9), shape=21)+
  facet_wrap(~phage)+
  theme_cowplot()+panel_border()



plot_grid(p1,p2,p3,p4, nrow = 2, labels = 'auto')

```

conclusions:  
a. model residuals are fairly close to normal distribution. Sign of good fit.  
b. The residuals are evenly distributed around 0, suggesting equal variance.  
c. Equal residual variance holds across time, though some difference seen in long-no phage.  
d. Equal residual variance holds across experimental units (flasks).

# post hoc  
What in the time X seed-bank interaction is causing a significant effect on population dynamics?  
We analyze the *estimated marginal means (EMM)*, following examples from: https://cran.r-project.org/web/packages/emmeans/vignettes/interactions.html

### Visualize EMM of interactions
```{r}
# plot 
emmip(m.best, time.fct~seed.bank)+
  theme_cowplot()
```

not clear from plot.


```{r}

coevo.emm <- emmeans(m.best, ~  seed.bank * time.fct)
coevo.pairs <- pairs(coevo.emm, simple="seed.bank")%>%
  tidy

p1 <- 
  coevo.pairs%>%
  # make time continuous for plotting.
  mutate(day=as.numeric(time.fct))%>%
  #plot
  ggplot(aes(x=day, y=estimate))+
    #add 0 line
  geom_hline(yintercept = 0, color="red")+
  geom_point()+
  geom_line()+
  facet_grid(contrast~.)+
  theme_cowplot()+
  panel_border()

p2 <- coevo.pairs%>%
  # focus on phage infected
  # filter(phage=="SPO1")%>%
  # # make time continuous for plotting.
  mutate(day=as.numeric(time.fct))%>%
  # arrange panel order
  mutate(contrast=fct_relevel(contrast, "long - short", after = 0))%>%
  #plot
  ggplot(aes(x=day, y=p.value))+
    #add 0.05 signifcance thrshold
  geom_hline(yintercept = 0.05, color="red")+
  geom_point()+
  geom_line()+
  facet_grid(contrast~.)+
  theme_cowplot()+
  panel_border()



plot_grid(p1,p2, labels = "auto") 
```

There are a few points that show significant difference between seed-bank treatments, but they are sporadically dispersed and are associated with different directions of change: three have higher spore percentage in short and 1 in long.

What about phage interaction with seed-bank?

```{r}
emm_sb.ph <- emmeans(m.best, pairwise ~ seed.bank | phage)

emm_sb.ph$contrasts

plot(emm_sb.ph)+
  theme_cowplot()+
  panel_border()

emmip(m.best, phage~seed.bank)+
  theme_cowplot()
```

The percentage of spores is higher in short seed bank treatment when infected by phage.

What is the trend over time?
```{r}

coevo.emm <- emmeans(m.best, ~ phage * time.fct)
coevo.pairs <- coevo.emm%>%
  tidy()

p1 <- 
  coevo.pairs%>%
  # make time continuous for plotting.
  mutate(day=as.numeric(time.fct))%>%
  #plot
  ggplot(aes(x=day, y=estimate))+
    #add trendline
  geom_smooth(method = "lm")+
  geom_point()+
  geom_line()+
  facet_grid(phage~.)+
  theme_cowplot()+
  panel_border()

p2 <- coevo.pairs%>%
  # make time continuous for plotting.
  mutate(day=as.numeric(time.fct))%>%
  #plot
  ggplot(aes(x=day, y=p.value))+
    #add 0.05 signifcance thrshold
  geom_hline(yintercept = 0.05, color="red")+
  geom_point()+
  geom_line()+
  facet_grid(phage~.)+
  theme_cowplot()+
  panel_border()



plot_grid(p1,p2, labels = "auto") 
```
A slight increase in percent of spores is seen in the phage infected cultures, an not in the absence of phage.

`Is there a way to estimate if this increase is significant`

# Summary