---
title: "RM-ANOVA for coevolution with a seed bank"
subtitle: "phenotypic evolution"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
library(renv)
# #install to project library the required packages
# renv::restore()
# # Initialize lock file when first setting repo
# renv::init()

library(here)
# # set root at project file ("coevolution-ts"). 
# # Done once at initial setup.
# here::set_here()

library(tidyverse)
library(cowplot)


library(nlme)
library(car)
# library(ggpubr)
library(broom)
library(broom.mixed)
library(emmeans)

# save the state of the project library to the lockfile (called renv.lock),
# renv::snapshot()

```

Load organized data  
```{r, echo=FALSE, message=FALSE}
d <- read_csv(here("data","resistance-pop-sum.csv"), )

```

### Adjusting variable types for lme

In the lme models used below the fixed effects need to be specified as factors. In this experiment these are: *seed.bank and time*. However for the auto correlation specification time needs to be specified as an integer. 

```{r}
d <- 
  d%>%
  rename(Time = t.host) %>% 
  # for the lme model all fixed effects need to be factors
  mutate(time.fct=as.factor(Time))%>%
  mutate(pop=as.factor(pop))%>%
  mutate(seed.bank=as.factor(seed.bank))%>%
  #ajust factor order for seed bank
  mutate(seed.bank = fct_rev(seed.bank))
```

Select response variable to be analyzed  

```{r}
var.response <- "pop_resistance" 

d <- d %>%
  mutate(response=pop_resistance ) %>% 
  mutate(flask = paste0(trt,"-", line)) 
```

Here we analyze **`r var.response` **.

```{r}
d%>%
ggplot(aes(x=Time, y=response))+
      geom_line(aes(group=flask,color=line), size=1, alpha=0.7)+
        facet_wrap(seed.bank~pop)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab(var.response)+
  ylim(0,NA)


```


### Test data for homogeneity of variances

This is an assumption of ANOVA tests.

Based on : https://www.datanovia.com/en/lessons/homogeneity-of-variance-test-in-r/
Using "Levene’s test" that according to website is the most commonly used test for this purpose. This test has a null hypothesis of equal variance. So getting P>0.05 suggests homogenic variance.  

```{r}
# Levene's test with multiple independent variables
car::leveneTest(response ~ pop*seed.bank*line, data = d)
```

**The data does not fulfill the assumption of equal variance across test groups.**


## Log transform the data  

Making it percent so that values remain positive

```{r}
d <- d%>%
  mutate(log.response=log10(100*response))

d%>%
ggplot(aes(x=Time, y=log.response,group=flask))+
      geom_line(aes(group=flask,color=line), size=1, alpha=0.7)+
        facet_wrap(seed.bank~pop)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab(paste(var.response,"(log10)"))


```


### Test transformed data for homogeneity of variances

```{r}
# Levene's test with multiple independent variables
car::leveneTest(log.response ~ pop * seed.bank * line, data = d)
```

**The data still does not fulfill the assumption of equal variance across test groups.**


# Box-Cox transformation

>`powerTransform` uses the maximum likelihood-like approach of Box and Cox (1964) to select a transformation of a univariate or multivariate response for normality, linearity and/or constant variance.  
>(help page for `car::powerTransform`)


```{r}
# Multivariate transformation to normality within levels of treatments
bx.cx <- powerTransform(response ~ pop*seed.bank*line, d)
summary(bx.cx)
```

*Tranformation is required, but not a simple log transformation*

Plot the profile log-likelihood for Box-Cox transformations.  
```{r}
  with(d, boxCox(response ~ pop*seed.bank*line, 
                 lambda = seq(-1, bx.cx$roundlam+3, by = 0.1),
                 family="bcPower"))
```

Transform using Box-Cox $\lambda$ (rounded). 
```{r}
d <- d%>%
  mutate(bxcx.response=bcPower(response, bx.cx$roundlam))

d%>%
ggplot(aes(x=Time, y=bxcx.response))+
      geom_line(aes(group=flask,color=line), size=1, alpha=0.7)+
        facet_wrap(pop~ seed.bank)+
      theme_bw()+
      panel_border()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab(paste(var.response,"(Box-Cox transformed)"))



```

### Test transformed data for homogeneity of variances

```{r}
# Levene's test with multiple independent variables
car::leveneTest(bxcx.response ~ pop*seed.bank*line, data = d)
```

**The data now fulfills the assumption of equal variance across test groups.**

#RM-ANOVA

##  vs. Spores in "with-seed-bank" populations

```{r}

cur.model <- d%>%
      # compare veg and spore in with-seed-bank
      filter(seed.bank == "with-seed-bank") %>% 
      lme(bxcx.response ~ pop * time.fct , 
          random = ~1|flask ,
          correlation=corARMA(form = ~ 1 | flask,  p = 1, q = 1),
            data = .)


anova(cur.model)
```

### Distribution of model residuals

Is the model any good?


```{r,fig.height=10, fig.width=15}

p1 <- 
  #qqplot by seed bank
  qqnorm(cur.model,~ resid(., type = "p")|pop, abline = c(0, 1))

p2 <-
  # standardized residuals versus fitted values by seed.bank
plot(cur.model, resid(., type = "p") ~ fitted(.) | pop, abline = 0)

p3 <-
broom.mixed::augment(cur.model)%>%
  ggplot(aes(Time,.resid)) +
  geom_hline(yintercept = 0)+
  geom_line(aes(color=line))+
  facet_wrap(~pop)+
  theme_cowplot()+panel_border()

p4 <-
broom.mixed::augment(cur.model)%>%
  ggplot(aes(line,.resid)) +
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(fill=line),alpha=.5, position = position_dodge(width = .9))+
  geom_point(aes(color=line), position = position_dodge(width = .9), shape=21)+
  facet_wrap(~pop)+
  theme_cowplot()+panel_border()



plot_grid(p1,p2,p3,p4, nrow = 2, labels = 'auto')

```

ok (L3 not great...)


## Seed-bank effect on veg populations

```{r}
cur.model <- d%>%
      # compare veg and spore in with-seed-bank
      filter(pop == "VEG") %>% 
      lme(response ~ seed.bank * time.fct , 
          random = ~1|flask ,
          correlation=corARMA(form = ~ 1 | flask, p=1, q=1),
            data = .)

anova(cur.model)
```



### Distribution of model residuals

Is the model any good?
  

```{r,fig.height=10, fig.width=15}

p1 <- 
  #qqplot by seed bank
  qqnorm(cur.model,~ resid(., type = "p")|seed.bank, abline = c(0, 1))

p2 <-
  # standardized residuals versus fitted values by seed.bank
plot(cur.model, resid(., type = "p") ~ fitted(.) | seed.bank, abline = 0)

p3 <-
broom.mixed::augment(cur.model)%>%
  ggplot(aes(Time,.resid)) +
  geom_hline(yintercept = 0)+
  geom_line(aes(color=line))+
  facet_wrap(~seed.bank)+
  theme_cowplot()+panel_border()

p4 <-
broom.mixed::augment(cur.model)%>%
  ggplot(aes(pop,.resid)) +
  geom_hline(yintercept = 0)+
  geom_boxplot(aes(fill=line),alpha=.5, position = position_dodge(width = .9))+
  geom_point(aes(color=line), position = position_dodge(width = .9), shape=21)+
  facet_wrap(~seed.bank)+
  theme_cowplot()+panel_border()



plot_grid(p1,p2,p3,p4, nrow = 2, labels = 'auto')

```


ok
