---
title: "RM-ANOVA for coevolution with a seed bank"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(renv)
# install to projectlibray the resuired packages
renv::restore()
# # Initialize lock file when first setting repo
# renv::init()

library(here)
# # set root at project file ("coevolution-ts"). 
# # Done once at initial setup.
# here::set_here()

library(tidyverse)
library(cowplot)


library(nlme)
library(pander)
library(emmeans)
library(codyn)
library(smooth)

# save the state of the project library to the lockfile (called renv.lock),
renv::snapshot()

```

## Population density data  
I have previously consolidated all the data on population density collected for this experiment. This includes:

| Population | What it is | Method of measurement |
| - | - | - |
|spore / mL | host spore density | flow-cytometry: low SYBR population |
|veg / mL | host vegetative cell density | flow-cytometry: high SYBR population |
| cell /mL | total host cell density (spore+veg) | flow-cytometry
| phage/ mL | total SPO1 phage density (free phage) | qPCR (SYBR)  with phage specific primers and lysate serial dilution as standard | 

```{r load data, echo=FALSE, message=FALSE}
d <- read_csv(here("data","coevolution_pop_density.csv"), )

#add columns on host strain and phage treatment  as encoded in culture name
d <- 
  d%>%
  mutate(host=case_when(grepl("W", culture) ~ "WT",
                         grepl("SN", culture) ~ "dSpoIIE",
                         grepl("dS", culture) ~ "dSpoIIE"))%>%
  
  mutate(phage=case_when(grepl("O", culture) ~ "SPO1",
                         grepl("Ct", culture) ~ "no_Phage",
                         grepl("ANC", line) ~ "no_Phage"))%>%

   mutate(seed.bank = fct_relevel(seed.bank, "long","short","none"))%>%
  #add flask to be used as grouping variable
  mutate(flask=paste(culture,line, sep = "-"))%>%
  #remove setup data
  filter(Time>=0)

# plot raw data
d%>%
    pivot_longer(cols=contains(".ml"), names_to="pop",values_to="num.ml")%>%
    ggplot(aes(x=Time, y=num.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      facet_grid(pop~phage)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 1e10))
```

Notably, in the first few days of the experiment cell numbers dropped below detection limit in 2 of 3 SNO lines. Missing data cannot be analyzed in RM-ANOVA. So for cell data let's look at data from day 4 and onward:

```{r}
d <- d%>%
  filter(Time>=4)

# plot total cells of raw data
d%>%
    ggplot(aes(x=Time, y=cell.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      facet_wrap(~phage)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 1e10))
```



## Model selection  

### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r arma, message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the sappce in which the model conerges is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
    #this fllowing combination does not work: "Coefficient matrix not invertible"
    if(P==1 & Q==1) next
    if(P==1 & Q==2) next

    cur.model <- 
        d%>%  
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

There are 4 models with lowest AIC, those having p=2 or p=6, and withe q=1 or q=2. Over all the q=1 models seem to be doin better.

### compare AR1 with ARMA
```{r}

# corAR1
cell.rm.AR1 <- 
  d%>%
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * Time , random = ~1|flask, 
            correlation = corAR1(form = ~ 1 | flask),
            data = .)

# corARMA models
cell.rm.ARMA.p2q1 <- 
        d%>%  
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=2,q=1),
            data = .)

cell.rm.ARMA.p2q2 <- 
        d%>%  
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=2,q=2),
            data = .)

cell.rm.ARMA.p6q1 <-
        d%>%
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * Time , random = ~1|flask,
            correlation = corARMA(form = ~ 1 | flask, p=6,q=1),
            data = .)

cell.rm.ARMA.p6q2 <-
        d%>%
  # looking only at total cells (cell.ml)
    lme(cell.ml ~ phage * seed.bank * Time , random = ~1|flask,
            correlation = corARMA(form = ~ 1 | flask, p=6,q=2),
            data = .)

AIC(cell.rm.AR1,cell.rm.ARMA.p2q1,cell.rm.ARMA.p2q2,cell.rm.ARMA.p6q1,cell.rm.ARMA.p6q2)%>%
  arrange(AIC)
```

The model ARMA p2q1 model is the best.

Is the improvment significant?  
```{r}
# AIC(cell.rm.AR1,cell.rm.ARMA.p2q1,cell.rm.ARMA.p2q2,cell.rm.ARMA.p6q1,cell.rm.ARMA.p6q2)%>%
#   arrange(desc(AIC))%>%
#   row.names()

anova(cell.rm.AR1,cell.rm.ARMA.p6q2,cell.rm.ARMA.p2q2,cell.rm.ARMA.p6q1,cell.rm.ARMA.p2q1)

```

Both q1 modes are equal, but improved over q2 models. All ARMA models seem better than te AR1 model. 

Selecting the ARMA model woth smaller degrees of freedom: p2q1.


### ARMA.p2q1  
```{r}
car::Anova(cell.rm.ARMA.p2q1)
```

Hoe sensetive would the result be to model selected?

### ARMA.p6q1  
```{r}
car::Anova(cell.rm.ARMA.p6q2)
```

### ARMA.p2q2  
```{r}
car::Anova(cell.rm.ARMA.p2q2)
```
### ARMA.p6q2  
```{r}
car::Anova(cell.rm.ARMA.p6q2)
```

### AR1  
```{r}
car::Anova(cell.rm.AR1)
```

All models show the same story:  
* The interaction of phage and time has the most sinificant influence on cell density.  
* Seed bank is not significant, but woud be a significant main effect at p-value of 0.1. 


## Post hoc

comparing means across seed bank for phage treatments seperatly.
First with phage: 
```{r, warning=F, message=F}

# I wrote a function to nicely plot the results of this test
source(here("code","functions", "plot_TukeyHSD.R"))

tuk <- TukeyHSD(aov(cell.ml~seed.bank,data = filter(d, phage=="SPO1")))

plot_TukeyHSD(tuk)

```

Without phage:
```{r}
tuk <- TukeyHSD(aov(cell.ml~seed.bank,data = filter(d, phage!="SPO1")))

plot_TukeyHSD(tuk)


```


The differences in cell density are only in phage treatment between no seed-bank and the other 2 seed-bank trearments.


# phage density

```{r}
d%>%
  filter(phage=="SPO1")%>%
    ggplot(aes(x=Time, y=phage.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      theme_bw()+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 1e10))
```


## Model selection  

### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r arma phage, message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the sappce in which the model conerges is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
    #this following combination does not work: "Citeration limit reached without convergence"
    if(P==4 & Q==1) next
    if(P==6 & Q==2) next
    #this following combination does not work: " Coefficient matrix not invertible"
    if(P==2 & Q==2) next

    cur.model <- 
        d%>%  
      filter(phage=="SPO1")%>%
  # looking only at total phage (phage.ml)
    lme(phage.ml ~ seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

There are 4 models with lowest AIC, those having (p=3 ,q=2 ); (p=4 ,q=2 );(p=5 ,q=1 )

### compare AR1 with ARMA
```{r}

# corAR1
phage.rm.AR1 <- 
  d%>%
        filter(phage=="SPO1")%>%
  # looking only at total phage (phage.ml)
    lme(phage.ml ~ seed.bank * Time , random = ~1|flask, 
            correlation = corAR1(form = ~ 1 | flask),
            data = .)

# corARMA models
phage.rm.ARMA.p3q2 <- 
        d%>%  
       filter(phage=="SPO1")%>%
  # looking only at total phage (phage.ml)
    lme(phage.ml ~ seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=3,q=2),
            data = .)

phage.rm.ARMA.p4q2 <- 
        d%>%  
       filter(phage=="SPO1")%>%
  # looking only at total phage (phage.ml)
    lme(phage.ml ~ seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=4,q=2),
            data = .)

phage.rm.ARMA.p5q1 <- 
        d%>%  
       filter(phage=="SPO1")%>%
  # looking only at total phage (phage.ml)
    lme(phage.ml ~ seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=5,q=1),
            data = .)

phage.rm.ARMA.p6q1 <- 
        d%>%  
       filter(phage=="SPO1")%>%
  # looking only at total phage (phage.ml)
    lme(phage.ml ~ seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=6,q=1),
            data = .)

AIC(phage.rm.AR1,phage.rm.ARMA.p3q2,phage.rm.ARMA.p4q2,phage.rm.ARMA.p5q1,phage.rm.ARMA.p6q1)%>%
  arrange(AIC)
```

The model ARMA p4q2 model is the best.

Is the improvment significant?  
```{r}
AIC(phage.rm.AR1,phage.rm.ARMA.p3q2,phage.rm.ARMA.p4q2,phage.rm.ARMA.p5q1,phage.rm.ARMA.p6q1)%>%
  arrange(desc(AIC))%>%
  row.names()

anova(phage.rm.AR1,phage.rm.ARMA.p6q1,phage.rm.ARMA.p3q2,phage.rm.ARMA.p5q1,phage.rm.ARMA.p4q2)

```

All ARMA modes are equal. All ARMA models seem better than te AR1 model. 

Selecting the ARMA model woth smaller degrees of freedom: p3q2


### ARMA.p3q2  
```{r}
car::Anova(phage.rm.ARMA.p3q2)
```

How sensetive would the result be to model selected?

### ARMA.p4q2  
```{r}
car::Anova(phage.rm.ARMA.p4q2)
```

### ARMA.p5q1 
```{r}
car::Anova(phage.rm.ARMA.p5q1)
```

### ARMA.p6q2 1
```{r}
car::Anova(phage.rm.ARMA.p6q1)
```

### AR1  
```{r}
car::Anova(phage.rm.AR1)
```

All models show the same story:  
Both phage and time have a significant influence on phage density. But not their interaction.


## Post hoc

comparing means across seed bank: 
```{r, warning=F, message=F}


tuk <- TukeyHSD(aov(phage.ml~seed.bank,data = filter(d, phage=="SPO1")))

plot_TukeyHSD(tuk)




```


# spore density

```{r}
d%>%
  filter(host=="WT")%>%
    ggplot(aes(x=Time, y=spore.ml))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      theme_bw()+
      facet_wrap(~phage)+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")+
      coord_cartesian(ylim = c(1e5, 1e9))
```


## Model selection  

### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r arma spore, message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the sappce in which the model conerges is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
    # #this following combination does not work: "Citeration limit reached without convergence"
    if(P==8 & Q==1) next
    if(P==3 & Q==2) next
    if(P>=5 & Q==2) next


    cur.model <- 
        d%>%  
      filter(host=="WT")%>%
  # looking only at total phage (phage.ml)
    lme(spore.ml ~ phage*seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

There are 2 models with lowest AIC, those having (p=7 ,q=0 ); (p=7 ,q=1 ).

### compare AR1 with ARMA
```{r}

# corAR1
spore.rm.AR1 <- 
          d%>%  
      filter(host=="WT")%>%
  # looking only at total phage (phage.ml)
    lme(spore.ml ~ phage*seed.bank * Time , random = ~1|flask, 
            correlation = corAR1(form = ~ 1 | flask),
            data = .)

# corARMA models
spore.rm.ARMA.p7q0 <- 
        d%>%  
      filter(host=="WT")%>%
  # looking only at total phage (phage.ml)
    lme(spore.ml ~ phage*seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=7,q=0),
            data = .)

spore.rm.ARMA.p7q1 <- 
        d%>%  
      filter(host=="WT")%>%
  # looking only at total phage (phage.ml)
    lme(spore.ml ~ phage*seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=7,q=1),
            data = .)
#adding a p2 model because of transfer regime
spore.rm.ARMA.p2q1 <- 
        d%>%  
      filter(host=="WT")%>%
  # looking only at total phage (phage.ml)
    lme(spore.ml ~ phage*seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=2,q=1),
            data = .)


AIC(spore.rm.AR1,spore.rm.ARMA.p7q0,spore.rm.ARMA.p7q1,spore.rm.ARMA.p2q1)%>%
  arrange(AIC)
```

```{r}
# AIC(spore.rm.AR1,spore.rm.ARMA.p7q0,spore.rm.ARMA.p7q1)%>%
#   arrange(desc(AIC))%>%
#   row.names()

anova(spore.rm.AR1,spore.rm.ARMA.p7q1,spore.rm.ARMA.p2q1,	spore.rm.ARMA.p7q0)
```

ARMA models are signficantly better

```{r}
anova(spore.rm.ARMA.p7q0)
```

phage:seed.bank:Time is significant.

```{r}
anova(spore.rm.ARMA.p7q1)
```

```{r}
anova(spore.rm.ARMA.p2q1)
```

```{r}
anova(spore.rm.AR1)
```

Significance of 3-way interaction holds in all models. 


## Post hoc

comparing means across seed bank for phage treatments seperatly, forst with phage: 
```{r, warning=F, message=F}


tuk <- TukeyHSD(aov(spore.ml~seed.bank,data = filter(d,host=="WT" &phage=="SPO1")))
tuk
plot_TukeyHSD(tuk)
```

And now wothout phage
```{r, warning=F, message=F}
tuk <- TukeyHSD(aov(spore.ml~seed.bank,data = filter(d,host=="WT" &phage!="SPO1")))
tuk
plot_TukeyHSD(tuk)


```

There is a seed-bank treatment effect only in the absence of phage. The effect seems very small (conf. interval on dofference almost at 0)

=========================================
# spore percent

```{r}
d.perc <- 
  d%>%
  filter(host=="WT")%>%
  mutate(perc.spore=spore.ml/cell.ml)

d.perc%>%
    ggplot(aes(x=Time, y=perc.spore))+
      geom_line(aes(group=flask,color=seed.bank), size=1, alpha=0.7)+
      theme_bw()+
      facet_wrap(~phage)+
      panel_border()+
      scale_y_log10()+
      theme(legend.position = "bottom",
            text=element_text(size=14))+
      ylab("cell/ml (log)")
```


## Model selection  

### ARMA covariance

This is a combined  auto-regeressive model, AR(p) and moving average model, MA(q). I here take a model selection approach to choose these parameters (lowest AIC).
```{r arma perc.spore, message = FALSE, warning = FALSE}



# initalise empty table
pq.aic <- tibble(p=1, q=1,aic=1)%>%
  filter(p>1)
  
# Playing around with p&q I found the sappce in which the model conerges is limited
for(Q in c(0:2)){ 
  for (P in c(1:(10-2*Q))){ 
    
    # # #this following combination does not work: "Citeration limit reached without convergence"
    if(P>=9 & Q==0) next
    if(P>=8 & Q==1) next
    if(P>=4 & Q==2) next


    cur.model <- 
        d.perc%>%  
  # looking only at total phage (phage.ml)
    lme(perc.spore ~ phage*seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=P,q=Q),
            data = .)
    
    pq.aic <- 
      tibble(p=P, q=Q,aic=AIC(cur.model))%>%
      bind_rows(pq.aic,.)
    
  }
}

pq.aic%>%
  mutate(q=as.character(q))%>%
  ggplot(aes(p,aic))+
  geom_line(aes(color=q),size=1)+
  geom_point(aes(color=q),size=3)+
  theme_bw()+
  scale_x_continuous(breaks = 1:10)+
  ggtitle("correlation = corARMA(p,q)")


```

There are 2 models with lowest AIC, those having (p=2 ,q=2 ); (p=2 ,q=3 ).

### compare AR1 with ARMA
```{r}

# corAR1
pspore.rm.AR1 <- 
          d.perc%>%  
    lme(perc.spore ~ phage*seed.bank * Time , random = ~1|flask, 
            correlation = corAR1(form = ~ 1 | flask),
            data = .)

# corARMA models
pspore.rm.ARMA.p2q2 <- 
   d.perc%>%  
    lme(perc.spore ~ phage*seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=2,q=2),
            data = .)

pspore.rm.ARMA.p2q3 <- 
   d.perc%>%  
    lme(perc.spore ~ phage*seed.bank * Time , random = ~1|flask, 
            correlation = corARMA(form = ~ 1 | flask, p=2,q=3),
            data = .)


AIC(pspore.rm.AR1,pspore.rm.ARMA.p2q2,pspore.rm.ARMA.p2q3)%>%
  arrange(AIC)
```

```{r}
# AIC(pspore.rm.AR1,pspore.rm.ARMA.p2q2,pspore.rm.ARMA.p2q3)%>%
#   arrange(desc(AIC))%>%
#   row.names()

anova(pspore.rm.AR1,pspore.rm.ARMA.p2q3,pspore.rm.ARMA.p2q2)
```

ARMA models are signficantly better

```{r}
anova(pspore.rm.ARMA.p2q2)
```

phage:seed.bank:Time is significant.

```{r}
anova(pspore.rm.ARMA.p2q3)
```

```{r}
anova(pspore.rm.AR1)
```



## Post hoc

comparing means across seed bank for phage treatments seperatly, forst with phage: 
```{r, warning=F, message=F}


tuk <- TukeyHSD(aov(perc.spore~seed.bank,data = filter(d.perc , phage=="SPO1")))
tuk
plot_TukeyHSD(tuk)
```

And now wothout phage
```{r, warning=F, message=F}
tuk <- TukeyHSD(aov(perc.spore~seed.bank,data = filter(d.perc , phage!="SPO1")))
tuk
plot_TukeyHSD(tuk)


```

There is a seed-bank treatment effect only woth phage. 